---
title: "MElaS Analysis"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
# Analysis of MElaS

## Import Packages

```{r}
library(dplyr) # version 1.1.4
library(ggplot2) # version 3.5.2
library(ggpubr) # version 0.6.0
library(tidyr) # version 1.3.1
library(FSA) # version 0.10.0
library(purrr) # version 1.0.4
library(stringr) # version 1.5.1
library(viridis) # version 0.6.5
```

## Load data

```{r}
# data of every sample and metabolite's MElaS
data<-read.csv("All_MElaS_Data_0v3v12.csv")

data$infected_parasites=as.factor(data$infected_parasites)
data$organ=as.factor(data$organ)
data$parasite_species=as.factor(data$parasite_species)

# data from GNPS
molecules<-read.csv("DiseaseToleranceGNPSResults.csv")

# data from original feature table
metadata<-read.csv("DisTol_mergeddata.csv")
metadata<-na.omit(metadata)
ATTRIBUTE_infected_parasites<-as.factor(metadata$ATTRIBUTE_infected_parasites)
ATTRIBUTE_endpoint<-as.factor(metadata$ATTRIBUTE_endpoint)

metadata<-filter(metadata, ATTRIBUTE_endpoint!="24w", ) #remove 24 weeks
metadata<-filter(metadata, ATTRIBUTE_infected_parasites!='10000', ) # remove 10000 concentration

```

## Filter data and make categories

```{r}
# Make a new dataset that includes categorizes
dataWithMElaSTypes <- data %>%
  mutate(
    Type = case_when(
      (logFC_3v0 >= -0.3 & logFC_3v0 <= 0.3) & (logFC_12v3>=-0.3 & logFC_12v3 <= 0.3)  ~ "flat",
      (logFC_3v0 >= -0.3 & logFC_3v0 <= 0.3) & (logFC_12v3 > 0.3) ~ "flat_up",
      (logFC_3v0 >= -0.3 & logFC_3v0 <= 0.3) & (logFC_12v3 < -0.3) ~ "flat_down",
      (logFC_3v0 > 0.3) & (logFC_12v3>=-0.3 & logFC_12v3<=0.3) ~ "up_flat",
      (logFC_3v0 < -0.3) & (logFC_12v3>=-0.3 & logFC_12v3<=0.3) ~ "down_flat",
      (logFC_12v3 > 0.3 & logFC_3v0 > 0.3) ~ "up_up",
      (logFC_12v3 < -0.3 & logFC_3v0 < -0.3) ~ "down_down",
      (logFC_3v0 > 0.3) & (logFC_12v3 < -0.3) ~ "up_down",
      (logFC_3v0 < -0.3) & (logFC_12v3 > 0.3) ~ "down_up"
    )
  )
dataWithMElaSTypes$organ <- factor(dataWithMElaSTypes$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

# Make a new dataset that removes all metabolites that were categorized "flat"
noflatData<-filter(dataWithMElaSTypes, Type!="flat", )

```

## Comparison of categories

```{r}
# Figure 3: Comparison of MElaS of each category within all metabolites (all metabolites included)
dataWithMElaSTypes$Type <- factor(dataWithMElaSTypes$Type, levels=c("up_down", "down_up", "flat_up", "flat_down", "up_flat", "down_flat", "up_up", "down_down", "flat"))
boxplot_cat <- ggplot(dataWithMElaSTypes, aes(x=Type, y=MElaS))+ geom_boxplot() + labs(title = "Boxplot Comparing MelaS of Each Category", x = "Category", y = "MElaS") +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(title=element_blank())
print(boxplot_cat)
plotname <- "thesis_pres_images/defense_images/boxplot_cat_MElaSValues.png"
ggsave(plotname, dpi = 300)

# Figure 4A: Bar plot illustrating distribution of each category within metabolites.Metabolites considered flat were included in this figure
barplot_cat_flat <- ggplot(data=dataWithMElaSTypes, aes(x = Type)) + geom_bar() + labs(x="Category", y="Count") +theme_bw() +theme(axis.text.y = element_text(size=13)) + theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=13)) + theme(title=element_blank())
print(barplot_cat_flat)
plotname <- "thesis_pres_images/defense_images/barplot_cat_count_flat.png"
ggsave(plotname, dpi = 300)

# Figure 4B: Bar plot illustrating distribution of each category within all metabolites. Metabolites considered flat were removed from this data set
dataWithMElaSTypesnoflat<-filter(dataWithMElaSTypes, Type!="flat", )
noflatData$Type <- factor(noflatData$Type, levels=c("up_down", "down_up", "flat_up", "flat_down", "up_flat", "down_flat", "up_up", "down_down"))

barplot_cat_noflat <- ggplot(noflatData, aes(x = Type)) + geom_bar() + labs(x="Category", y="Count") +theme_bw() +theme(axis.text.y = element_text(size=13)) + theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=13)) + theme(title=element_blank())
print(barplot_cat_noflat)
plotname <- "thesis_pres_images/defense_images/barplot_cat_count_noflat.png"
ggsave(plotname, dpi = 300)
```

## Chemical annotation of all metabolites and most and least elastic metabolites

```{r}
#analysis of overlap of metabolites shared among different organs
noflatData_overlapping <- noflatData %>%
  group_by(ID) %>%
  mutate(count_occurrences = n()) %>%
  ungroup() 

# Merge GNPS data with MElaS data to idenitfy annoted molecules
dataWithMElaSTypes <- dataWithMElaSTypes %>%
  mutate(extracted_ID = str_remove_all(ID, "X(\\d+).(\\d+)_(\\d+).(\\d+)_"))

merged_all_df <- merge(
  molecules,
  dataWithMElaSTypes,
  by.x = "X.Scan.",
  by.y = "extracted_ID",
  all = FALSE 
)

merged_all_df <- merged_all_df[, c( 'X.Scan.', 'superclass', 'class', 'subclass', 'ID', 'MElaS', 'logFC_3v0', 'logFC_12v3', 'FC_0w', 'FC_3w', 'FC_12w', 'FC_24w', 'organ', 'parasite_species', 'infected_parasites', 'Type')]

# Identify annotated molecules for most elastic molecules in each organ
all_top_df <- data.frame()
for (organ in unique(merged_all_df$organ)) {
  temp <- merged_all_df[merged_all_df$organ == organ,]
  
  top<-top_n(temp, 50, MElaS) #metabolites with top MElaS scores
  all_top_df <- rbind(all_top_df, top)
}
#write.csv(all_top_df, "all_top_df.csv", row.names = FALSE)

# Identify annotated molecules for least elastic molecules in each organ
all_bottom_df <- data.frame()
for (organ in unique(merged_all_df$organ)) {
  temp <- merged_all_df[merged_all_df$organ == organ,]
  
  bottom<-top_n(temp, -50, MElaS) #metabolites with lowest MElaS scores
  bottom <- bottom[sample(nrow(bottom), 50), ]
  all_bottom_df <- rbind(all_bottom_df, bottom)
}
#write.csv(all_bottom_df, "all_bottom_df.csv", row.names = FALSE)
```

# Analysis of most and least elastic metabolite per parasite strain and inoculation concentration

```{r}
# Make data frames for each strain and concentration combination
JR_1000_noflatData<- noflatData %>%
  filter(parasite_species=="JR", infected_parasites=="1000")

JR_100_noflatData<- noflatData %>%
  filter(parasite_species=="JR", infected_parasites=="100")

CL_1000_noflatData<- noflatData %>%
  filter(parasite_species=="CL", infected_parasites=="1000")

CL_100_noflatData<- noflatData %>%
  filter(parasite_species=="CL", infected_parasites=="100")
```

#JR, 1000 Analysis
## Top 10 MElaS per each organ
```{r}
noflatDatalist = list()
for (organ in unique(JR_1000_noflatData$organ)) {
  temp <- JR_1000_noflatData[JR_1000_noflatData$organ == organ,]
  
  top<-top_n(temp, 10, MElaS) #metabolites with top MElaS scores
  
  long_noflatData <- pivot_longer(top, cols = c(logFC_3v0, logFC_12v3),names_to = "Ratio", values_to = "LogFC")
  long_noflatData$Ratio <- gsub("Ratio", "", long_noflatData$Ratio)
  noflatDatalist[[length(noflatDatalist) + 1]] <- top
}
all_top_MElaSJR1000_df<-do.call(rbind, noflatDatalist)
```

## Bottom 10 MElaS per each organ

```{r}
noflatDatalist2 = list()
for (organ in unique(JR_1000_noflatData$organ)) {
  temp <- JR_1000_noflatData[JR_1000_noflatData$organ == organ,]
  
  bottom<-top_n(temp, -10, MElaS) #metabolites with lowest MElaS scores
  bottom <- bottom[sample(nrow(bottom), 10), ] #randomly sample 10
  long_noflatData <- pivot_longer(bottom, cols = c(logFC_3v0, logFC_12v3), 
                          names_to = "Ratio", values_to = "LogFC")
  long_noflatData$Ratio <- gsub("Ratio", "", long_noflatData$Ratio)
  noflatDatalist2[[length(noflatDatalist2) + 1]] <- bottom

}
all_bottom_MElaSJR1000_df<-do.call(rbind, noflatDatalist2)
```

## Boxplot comparing peak areas from feature table of each top MElaS metabolites by post infection days and condition

```{r}
JR_1000_metadata <- metadata %>%
  filter(ATTRIBUTE_parasite_species!="CL_Luc", ATTRIBUTE_infected_parasites!="100") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='1000' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

```

## Top metabolite boxplots (only in Heart D)

```{r}
all_top_MElaSJR1000_df_Heart_D<-filter(all_top_MElaSJR1000_df, organ=="Heart_D") # can change to other organs of interest
top_metabolite<-all_top_MElaSJR1000_df_Heart_D$ID 
column_list <- list()
for (name in top_metabolite) {
  column_list[[name]] <- metadata[[name]] 
} #make a dataset that only includes metabolites with top MElas scores

new_df <- as.data.frame(column_list)
new_df<-cbind(new_df, metadata[c(1,2,3, 4, 5)]) #make new data set with data from feature table of top 10 metabolites

new_df <- new_df %>%
  filter(ATTRIBUTE_parasite_species!="CL_Luc", ATTRIBUTE_infected_parasites!="100") %>%
  filter(ATTRIBUTE_organ=="Heart_D") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='1000' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

# boxplot comparing peak areas of each top MElaS metabolites by post infection days and condition. A box plot is produced for each metabolite 
for (metabolite in top_metabolite) {
  print(metabolite)
  plot<-ggplot(new_df, aes(x = as.factor(ATTRIBUTE_endpoint), y = (.data[[metabolite]]), fill = as.factor(ATTRIBUTE_infected_parasites))) +
  geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
  labs(title=paste0("Peak Area of Top MElaS ", metabolite), x = "Endpoint (weeks)", y = "Peak Area", fill="Infection Status") + geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9) +theme_bw() + scale_fill_manual(values = c("infected" = "red", "uninfected" = "grey")) +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(legend.text=element_text(size=13))+  theme(legend.title=element_text(size=rel(1.8)))
  print(plot)
  #plotname <- paste0("images2/top_metabolites_plots/", metabolite, "boxplot.png")
  #ggsave(plotname, dpi = 300)
  
}
```

## Bottom metabolite boxplots (only in Heart D)

```{r}
all_bottom_MElaSJR1000_df_Heart_D<-filter(all_bottom_MElaSJR1000_df, organ=="Heart_D")
bottom_metabolite<-all_bottom_MElaSJR1000_df_Heart_D$ID 
column_list <- list()
for (name in bottom_metabolite) {
  column_list[[name]] <- metadata[[name]] 
} 

new_df <- as.data.frame(column_list)
new_df<-cbind(new_df, metadata[c(1,2,3, 4, 5)])

new_df <- new_df %>%
  filter(ATTRIBUTE_parasite_species!="CL_Luc", ATTRIBUTE_infected_parasites!="100") %>%
  filter(ATTRIBUTE_organ=="Heart_D") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='1000' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

#boxplot comparing peak areas of each top MElaS metabolites by post infection days and condition. A box plot is produced for each metabolite 
for (metabolite in bottom_metabolite) {
  print(metabolite)
  
  plot<-ggplot(new_df, aes(x = as.factor(ATTRIBUTE_endpoint), y = (.data[[metabolite]]), fill = as.factor(ATTRIBUTE_infected_parasites))) +
  geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
  labs(title=paste0("Peak Area of Bottom MElaS ", metabolite), x = "Endpoint (weeks)", y = "Peak Area", fill="Infection Status") + geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9) +theme_bw() + scale_fill_manual(values = c("infected" = "red", "uninfected" = "grey")) +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(legend.text=element_text(size=13))+  theme(legend.title=element_text(size=rel(1.8)))
  print(plot)
  #plotname <- paste0("images2/bottom_metabolites_plots/", metabolite, "boxplot_bottom.png")
  #ggsave(plotname, dpi = 300)
}
```

## Plot lineplots of top 10 Metabolites and there FC at FC_0w, FC_3w and LogFC12w

```{r}
all_top_MElaSJR1000_df<-filter(all_top_MElaSJR1000_df, organ=="Heart_D")
for (organ in unique(all_top_MElaSJR1000_df$organ)) {
  temp <- all_top_MElaSJR1000_df[all_top_MElaSJR1000_df$organ == organ,]
  
  long_noflatData_with_time_top <- pivot_longer(temp, cols = c(FC_0w, FC_3w, FC_12w), 
                          names_to = "Time", values_to = "LogFC")
  long_noflatData_with_time_top$Time <- factor(long_noflatData_with_time_top$Time, levels=c("FC_0w", "FC_3w", "FC_12w"))
  
  for (metabolite in unique(long_noflatData_with_time_top$ID)) {
    temp<-long_noflatData_with_time_top[long_noflatData_with_time_top$ID==metabolite, ]
    p=ggplot(temp, aes(x = Time, y = LogFC, group=1)) + geom_line() + geom_point() + labs(title = paste0("Top MElas: ", metabolite, " FC Between Endpoints in ", organ), x = "Endpoint (weeks)", y = "FC")+ theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(2.8))) + theme(axis.title.x = element_text(size = rel(1.8))) +theme(axis.title.x = element_blank(), axis.text.x = element_blank(), title= element_blank())
  print(p)
  #plotname2 <- paste0("images2/top_metabolites_plots/", metabolite, "lineplot.png")
  #ggsave(plotname2, dpi = 300) 

  }
}
```

## Plot lineplots of bottom 10 Metabolites and there FC at FC_0w, FC_3w and LogFC12w

```{r}
all_bottom_MElaSJR1000_df<-filter(all_bottom_MElaSJR1000_df, organ=="Heart_D")
for (organ in unique(all_bottom_MElaSJR1000_df$organ)) {
  temp <- all_bottom_MElaSJR1000_df[all_bottom_MElaSJR1000_df$organ == organ,]
  
  long_noflatData_with_time_bottom <- pivot_longer(temp, cols = c(FC_0w, FC_3w, FC_12w), 
                          names_to = "Time", values_to = "LogFC")
  long_noflatData_with_time_bottom$Time <- factor(long_noflatData_with_time_bottom$Time, levels=c("FC_0w", "FC_3w", "FC_12w"))
  
  for (metabolite in unique(long_noflatData_with_time_bottom$ID)) {
    temp<-long_noflatData_with_time_bottom[long_noflatData_with_time_bottom$ID==metabolite, ]
    p=ggplot(temp, aes(x = Time, y = LogFC, group=1)) + geom_line() + geom_point() + labs(title = paste0("Bottom MElas: ", metabolite, " FC Between Endpoints in ", organ), x = "Endpoint Rations", y = "FC")+ theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(2.8))) + theme(axis.title.x = element_text(size = rel(1.8))) +theme(axis.title.x = element_blank(), axis.text.x = element_blank(), title= element_blank())
  print(p)
  #plotname2 <- paste0("images2/", metabolite, "lineplot_bottom.png")
  #ggsave(plotname2, dpi = 300)
  }
}

```

#JR, 100 Analysis
## Top 10 MElaS per each organ (work in progess)
```{r}
noflatDatalist = list()
for (organ in unique(JR_100_noflatData$organ)) {
  temp <- JR_100_noflatData[JR_100_noflatData$organ == organ,]
  
  top<-top_n(temp, 10, MElaS) #metabolites with top MElaS scores
  
  long_noflatData <- pivot_longer(top, cols = c(logFC_3v0, logFC_12v3),names_to = "Ratio", values_to = "LogFC")
  long_noflatData$Ratio <- gsub("Ratio", "", long_noflatData$Ratio)
  noflatDatalist[[length(noflatDatalist) + 1]] <- top
}
all_top_MElaSJR1000_df<-do.call(rbind, noflatDatalist)
```

## Bottom 10 MElaS per each organ

```{r}
noflatDatalist2 = list()
for (organ in unique(JR_100_noflatData$organ)) {
  temp <- JR_100_noflatData[JR_100_noflatData$organ == organ,]
  
  bottom<-top_n(temp, -10, MElaS) #metabolites with lowest MElaS scores
  bottom <- bottom[sample(nrow(bottom), 10), ] #randomly sample 10
  long_noflatData <- pivot_longer(bottom, cols = c(logFC_3v0, logFC_12v3), 
                          names_to = "Ratio", values_to = "LogFC")
  long_noflatData$Ratio <- gsub("Ratio", "", long_noflatData$Ratio)
  noflatDatalist2[[length(noflatDatalist2) + 1]] <- bottom

}
all_bottom_MElaSJR1000_df<-do.call(rbind, noflatDatalist2)
```

## Boxplot comparing peak areas from feature table of each top MElaS metabolites by post infection days and condition

```{r}
JR_100_metadata <- metadata %>%
  filter(ATTRIBUTE_parasite_species!="CL_Luc", ATTRIBUTE_infected_parasites!="100") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='1000' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

```

## Top metabolite boxplots (only in Heart D)

```{r}
all_top_MElaSJR1000_df_Heart_D<-filter(all_top_MElaSJR1000_df, organ=="Heart_D") # can change to other organs of interest
top_metabolite<-all_top_MElaSJR1000_df_Heart_D$ID 
column_list <- list()
for (name in top_metabolite) {
  column_list[[name]] <- metadata[[name]] 
} #make a dataset that only includes metabolites with top MElas scores

new_df <- as.data.frame(column_list)
new_df<-cbind(new_df, metadata[c(1,2,3, 4, 5)]) #make new data set with data from feature table of top 10 metabolites

new_df <- new_df %>%
  filter(ATTRIBUTE_parasite_species!="CL_Luc", ATTRIBUTE_infected_parasites!="100") %>%
  filter(ATTRIBUTE_organ=="Heart_D") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='1000' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

# boxplot comparing peak areas of each top MElaS metabolites by post infection days and condition. A box plot is produced for each metabolite 
for (metabolite in top_metabolite) {
  print(metabolite)
  plot<-ggplot(new_df, aes(x = as.factor(ATTRIBUTE_endpoint), y = (.data[[metabolite]]), fill = as.factor(ATTRIBUTE_infected_parasites))) +
  geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
  labs(title=paste0("Peak Area of Top MElaS ", metabolite), x = "Endpoint (weeks)", y = "Peak Area", fill="Infection Status") + geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9) +theme_bw() + scale_fill_manual(values = c("infected" = "red", "uninfected" = "grey")) +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(legend.text=element_text(size=13))+  theme(legend.title=element_text(size=rel(1.8)))
  print(plot)
  plotname <- paste0("images2/top_metabolites_plots/", metabolite, "boxplot.png")
  ggsave(plotname, dpi = 300)
  
}
```

## Bottom metabolite boxplots (only in Heart D)

```{r}
all_bottom_MElaSJR1000_df_Heart_D<-filter(all_bottom_MElaSJR1000_df, organ=="Heart_D")
bottom_metabolite<-all_bottom_MElaSJR1000_df_Heart_D$ID 
column_list <- list()
for (name in bottom_metabolite) {
  column_list[[name]] <- metadata[[name]] 
} 

new_df <- as.data.frame(column_list)
new_df<-cbind(new_df, metadata[c(1,2,3, 4, 5)])

new_df <- new_df %>%
  filter(ATTRIBUTE_parasite_species!="CL_Luc", ATTRIBUTE_infected_parasites!="100") %>%
  filter(ATTRIBUTE_organ=="Heart_D") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='1000' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

#boxplot comparing peak areas of each top MElaS metabolites by post infection days and condition. A box plot is produced for each metabolite 
for (metabolite in bottom_metabolite) {
  print(metabolite)
  
  plot<-ggplot(new_df, aes(x = as.factor(ATTRIBUTE_endpoint), y = (.data[[metabolite]]), fill = as.factor(ATTRIBUTE_infected_parasites))) +
  geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
  labs(title=paste0("Peak Area of Bottom MElaS ", metabolite), x = "Endpoint (weeks)", y = "Peak Area", fill="Infection Status") + geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9) +theme_bw() + scale_fill_manual(values = c("infected" = "red", "uninfected" = "grey")) +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(legend.text=element_text(size=13))+  theme(legend.title=element_text(size=rel(1.8)))
  print(plot)
  plotname <- paste0("images2/bottom_metabolites_plots/", metabolite, "boxplot_bottom.png")
  ggsave(plotname, dpi = 300)
}
```

## Plot lineplots of top 10 Metabolites and there FC at FC_0w, FC_3w and LogFC12w

```{r}
all_top_MElaSJR1000_df<-filter(all_top_MElaSJR1000_df, organ=="Heart_D")
for (organ in unique(all_top_MElaSJR1000_df$organ)) {
  temp <- all_top_MElaSJR1000_df[all_top_MElaSJR1000_df$organ == organ,]
  
  long_noflatData_with_time_top <- pivot_longer(temp, cols = c(FC_0w, FC_3w, FC_12w), 
                          names_to = "Time", values_to = "LogFC")
  long_noflatData_with_time_top$Time <- factor(long_noflatData_with_time_top$Time, levels=c("FC_0w", "FC_3w", "FC_12w"))
  
  for (metabolite in unique(long_noflatData_with_time_top$ID)) {
    temp<-long_noflatData_with_time_top[long_noflatData_with_time_top$ID==metabolite, ]
    p=ggplot(temp, aes(x = Time, y = LogFC, group=1)) + geom_line() + geom_point() + labs(title = paste0("Top MElas: ", metabolite, " FC Between Endpoints in ", organ), x = "Endpoint (weeks)", y = "FC")+ theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(2.8))) + theme(axis.title.x = element_text(size = rel(1.8))) +theme(axis.title.x = element_blank(), axis.text.x = element_blank(), title= element_blank())
  print(p)
  plotname2 <- paste0("images2/top_metabolites_plots/", metabolite, "lineplot.png")
  ggsave(plotname2, dpi = 300) 

  }
}
```

## Plot lineplots of bottom 10 Metabolites and there FC at FC_0w, FC_3w and LogFC12w

```{r}
all_bottom_MElaSJR1000_df<-filter(all_bottom_MElaSJR1000_df, organ=="Heart_D")
for (organ in unique(all_bottom_MElaSJR1000_df$organ)) {
  temp <- all_bottom_MElaSJR1000_df[all_bottom_MElaSJR1000_df$organ == organ,]
  
  long_noflatData_with_time_bottom <- pivot_longer(temp, cols = c(FC_0w, FC_3w, FC_12w), 
                          names_to = "Time", values_to = "LogFC")
  long_noflatData_with_time_bottom$Time <- factor(long_noflatData_with_time_bottom$Time, levels=c("FC_0w", "FC_3w", "FC_12w"))
  
  for (metabolite in unique(long_noflatData_with_time_bottom$ID)) {
    temp<-long_noflatData_with_time_bottom[long_noflatData_with_time_bottom$ID==metabolite, ]
    p=ggplot(temp, aes(x = Time, y = LogFC, group=1)) + geom_line() + geom_point() + labs(title = paste0("Bottom MElas: ", metabolite, " FC Between Endpoints in ", organ), x = "Endpoint Rations", y = "FC")+ theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(2.8))) + theme(axis.title.x = element_text(size = rel(1.8))) +theme(axis.title.x = element_blank(), axis.text.x = element_blank(), title= element_blank())
  print(p)
  plotname2 <- paste0("images2/", metabolite, "lineplot_bottom.png")
  ggsave(plotname2, dpi = 300)
  }
}

```

# CL, 1000 Analysis
## Top 10 MElaS per each organ

```{r}
noflatDatalist = list()
for (organ in unique(CL_1000_noflatData$organ)) {
  temp <- CL_1000_noflatData[CL_1000_noflatData$organ == organ,]
  
  top<-top_n(temp, 10, MElaS) #metabolites with top MElaS scores
  
  long_noflatData <- pivot_longer(top, cols = c(logFC_3v0, logFC_12v3),names_to = "Ratio", values_to = "LogFC")
  long_noflatData$Ratio <- gsub("Ratio", "", long_noflatData$Ratio)
  noflatDatalist[[length(noflatDatalist) + 1]] <- top
}
all_top_MElaSJR1000_df<-do.call(rbind, noflatDatalist)
```

## Bottom 10 MElaS per each organ

```{r}
noflatDatalist2 = list()
for (organ in unique(CL_1000_noflatData$organ)) {
  temp <- CL_1000_noflatData[CL_1000_noflatData$organ == organ,]
  
  bottom<-top_n(temp, -10, MElaS) #metabolites with lowest MElaS scores
  bottom <- bottom[sample(nrow(bottom), 10), ] #randomly sample 10
  long_noflatData <- pivot_longer(bottom, cols = c(logFC_3v0, logFC_12v3), 
                          names_to = "Ratio", values_to = "LogFC")
  long_noflatData$Ratio <- gsub("Ratio", "", long_noflatData$Ratio)
  noflatDatalist2[[length(noflatDatalist2) + 1]] <- bottom

}
all_bottom_MElaSJR1000_df<-do.call(rbind, noflatDatalist2)
```

## Boxplot comparing peak areas from feature table of each top MElaS metabolites by post infection days and condition

```{r}
CL_1000_metadata <- metadata %>%
  filter(ATTRIBUTE_parasite_species!="JR_Luc", ATTRIBUTE_infected_parasites!="100") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='1000' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

```

## Top metabolite boxplots (only in Heart D)

```{r}
all_top_MElaSJR1000_df_Heart_D<-filter(all_top_MElaSJR1000_df, organ=="Heart_D") # can change to other organs of interest
top_metabolite<-all_top_MElaSJR1000_df_Heart_D$ID 
column_list <- list()
for (name in top_metabolite) {
  column_list[[name]] <- metadata[[name]] 
} #make a dataset that only includes metabolites with top MElas scores

new_df <- as.data.frame(column_list)
new_df<-cbind(new_df, metadata[c(1,2,3, 4, 5)]) #make new data set with data from feature table of top 10 metabolites

new_df <- new_df %>%
  filter(ATTRIBUTE_parasite_species!="JR_Luc", ATTRIBUTE_infected_parasites!="100") %>%
  filter(ATTRIBUTE_organ=="Heart_D") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='1000' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

# boxplot comparing peak areas of each top MElaS metabolites by post infection days and condition. A box plot is produced for each metabolite 
for (metabolite in top_metabolite) {
  print(metabolite)
  plot<-ggplot(new_df, aes(x = as.factor(ATTRIBUTE_endpoint), y = (.data[[metabolite]]), fill = as.factor(ATTRIBUTE_infected_parasites))) +
  geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
  labs(title=paste0("Peak Area of Top MElaS ", metabolite), x = "Endpoint (weeks)", y = "Peak Area", fill="Infection Status") + geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9) +theme_bw() + scale_fill_manual(values = c("infected" = "red", "uninfected" = "grey")) +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(legend.text=element_text(size=13))+  theme(legend.title=element_text(size=rel(1.8)))
  print(plot)
  plotname <- paste0("images2/top_metabolites_plots/", metabolite, "boxplot.png")
  ggsave(plotname, dpi = 300)
  
}
```

## Bottom metabolite boxplots (only in Heart D)

```{r}
all_bottom_MElaSJR1000_df_Heart_D<-filter(all_bottom_MElaSJR1000_df, organ=="Heart_D")
bottom_metabolite<-all_bottom_MElaSJR1000_df_Heart_D$ID 
column_list <- list()
for (name in bottom_metabolite) {
  column_list[[name]] <- metadata[[name]] 
} 

new_df <- as.data.frame(column_list)
new_df<-cbind(new_df, metadata[c(1,2,3, 4, 5)])

new_df <- new_df %>%
  filter(ATTRIBUTE_parasite_species!="JR_Luc", ATTRIBUTE_infected_parasites!="100") %>%
  filter(ATTRIBUTE_organ=="Heart_D") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='1000' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

#boxplot comparing peak areas of each top MElaS metabolites by post infection days and condition. A box plot is produced for each metabolite 
for (metabolite in bottom_metabolite) {
  print(metabolite)
  
  plot<-ggplot(new_df, aes(x = as.factor(ATTRIBUTE_endpoint), y = (.data[[metabolite]]), fill = as.factor(ATTRIBUTE_infected_parasites))) +
  geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
  labs(title=paste0("Peak Area of Bottom MElaS ", metabolite), x = "Endpoint (weeks)", y = "Peak Area", fill="Infection Status") + geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9) +theme_bw() + scale_fill_manual(values = c("infected" = "red", "uninfected" = "grey")) +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(legend.text=element_text(size=13))+  theme(legend.title=element_text(size=rel(1.8)))
  print(plot)
  plotname <- paste0("images2/bottom_metabolites_plots/", metabolite, "boxplot_bottom.png")
  ggsave(plotname, dpi = 300)
}
```

## Plot lineplots of top 10 Metabolites and there FC at FC_0w, FC_3w and LogFC12w

```{r}
all_top_MElaSJR1000_df<-filter(all_top_MElaSJR1000_df, organ=="Heart_D")
for (organ in unique(all_top_MElaSJR1000_df$organ)) {
  temp <- all_top_MElaSJR1000_df[all_top_MElaSJR1000_df$organ == organ,]
  
  long_noflatData_with_time_top <- pivot_longer(temp, cols = c(FC_0w, FC_3w, FC_12w), 
                          names_to = "Time", values_to = "LogFC")
  long_noflatData_with_time_top$Time <- factor(long_noflatData_with_time_top$Time, levels=c("FC_0w", "FC_3w", "FC_12w"))
  
  for (metabolite in unique(long_noflatData_with_time_top$ID)) {
    temp<-long_noflatData_with_time_top[long_noflatData_with_time_top$ID==metabolite, ]
    p=ggplot(temp, aes(x = Time, y = LogFC, group=1)) + geom_line() + geom_point() + labs(title = paste0("Top MElas: ", metabolite, " FC Between Endpoints in ", organ), x = "Endpoint (weeks)", y = "FC")+ theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(2.8))) + theme(axis.title.x = element_text(size = rel(1.8))) +theme(axis.title.x = element_blank(), axis.text.x = element_blank(), title= element_blank())
  print(p)
  plotname2 <- paste0("images2/top_metabolites_plots/", metabolite, "lineplot.png")
  ggsave(plotname2, dpi = 300) 

  }
}
```

## Plot lineplots of bottom 10 Metabolites and there FC at FC_0w, FC_3w and LogFC12w

```{r}
all_bottom_MElaSJR1000_df<-filter(all_bottom_MElaSJR1000_df, organ=="Heart_D")
for (organ in unique(all_bottom_MElaSJR1000_df$organ)) {
  temp <- all_bottom_MElaSJR1000_df[all_bottom_MElaSJR1000_df$organ == organ,]
  
  long_noflatData_with_time_bottom <- pivot_longer(temp, cols = c(FC_0w, FC_3w, FC_12w), 
                          names_to = "Time", values_to = "LogFC")
  long_noflatData_with_time_bottom$Time <- factor(long_noflatData_with_time_bottom$Time, levels=c("FC_0w", "FC_3w", "FC_12w"))
  
  for (metabolite in unique(long_noflatData_with_time_bottom$ID)) {
    temp<-long_noflatData_with_time_bottom[long_noflatData_with_time_bottom$ID==metabolite, ]
    p=ggplot(temp, aes(x = Time, y = LogFC, group=1)) + geom_line() + geom_point() + labs(title = paste0("Bottom MElas: ", metabolite, " FC Between Endpoints in ", organ), x = "Endpoint Rations", y = "FC")+ theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(2.8))) + theme(axis.title.x = element_text(size = rel(1.8))) +theme(axis.title.x = element_blank(), axis.text.x = element_blank(), title= element_blank())
  print(p)
  plotname2 <- paste0("images2/", metabolite, "lineplot_bottom.png")
  ggsave(plotname2, dpi = 300)
  }
}

```

# CL, 100 Analysis
## Top 10 MElaS per each organ
```{r}
noflatDatalist = list()
for (organ in unique(CL_100_noflatData$organ)) {
  temp <- CL_100_noflatData[CL_100_noflatData$organ == organ,]
  
  top<-top_n(temp, 10, MElaS) #metabolites with top MElaS scores
  
  long_noflatData <- pivot_longer(top, cols = c(logFC_3v0, logFC_12v3),names_to = "Ratio", values_to = "LogFC")
  long_noflatData$Ratio <- gsub("Ratio", "", long_noflatData$Ratio)
  noflatDatalist[[length(noflatDatalist) + 1]] <- top
}
all_top_MElaSCL100_df<-do.call(rbind, noflatDatalist)
```

## Bottom 10 MElaS per each organ

```{r}
noflatDatalist2 = list()
for (organ in unique(CL_100_noflatData$organ)) {
  temp <- CL_100_noflatData[CL_100_noflatData$organ == organ,]
  
  bottom<-top_n(temp, -10, MElaS) #metabolites with lowest MElaS scores
  bottom <- bottom[sample(nrow(bottom), 10), ] #randomly sample 10
  long_noflatData <- pivot_longer(bottom, cols = c(logFC_3v0, logFC_12v3), 
                          names_to = "Ratio", values_to = "LogFC")
  long_noflatData$Ratio <- gsub("Ratio", "", long_noflatData$Ratio)
  noflatDatalist2[[length(noflatDatalist2) + 1]] <- bottom

}
all_bottom_MElaSCL100_df<-do.call(rbind, noflatDatalist2)
```

## Boxplot comparing peak areas from feature table of each top MElaS metabolites by post infection days and condition

```{r}
CL_100_metadata <- metadata %>%
  filter(ATTRIBUTE_parasite_species!="JR_Luc", ATTRIBUTE_infected_parasites!="1000") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='100' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

```

## Top metabolite boxplots (only in Heart D)

```{r}
all_top_MElaSCL100_df_Heart_D<-filter(all_top_MElaSCL100_df, organ=="Heart_D") # can change to other organs of interest
top_metabolite<-all_top_MElaSCL100_df_Heart_D$ID 
column_list <- list()
for (name in top_metabolite) {
  column_list[[name]] <- metadata[[name]] 
} #make a dataset that only includes metabolites with top MElas scores

new_df <- as.data.frame(column_list)
new_df<-cbind(new_df, metadata[c(1,2,3, 4, 5)]) #make new data set with data from feature table of top 10 metabolites

new_df <- new_df %>%
  filter(ATTRIBUTE_parasite_species!="CL_Luc", ATTRIBUTE_infected_parasites!="100") %>%
  filter(ATTRIBUTE_organ=="Heart_D") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='1000' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

# boxplot comparing peak areas of each top MElaS metabolites by post infection days and condition. A box plot is produced for each metabolite 
for (metabolite in top_metabolite) {
  print(metabolite)
  plot<-ggplot(new_df, aes(x = as.factor(ATTRIBUTE_endpoint), y = (.data[[metabolite]]), fill = as.factor(ATTRIBUTE_infected_parasites))) +
  geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
  labs(title=paste0("Peak Area of Top MElaS ", metabolite), x = "Endpoint (weeks)", y = "Peak Area", fill="Infection Status") + geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9) +theme_bw() + scale_fill_manual(values = c("infected" = "red", "uninfected" = "grey")) +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(legend.text=element_text(size=13))+  theme(legend.title=element_text(size=rel(1.8)))
  print(plot)
  plotname <- paste0("images2/top_metabolites_plots/", metabolite, "boxplot.png")
  ggsave(plotname, dpi = 300)
  
}
```

## Bottom metabolite boxplots (only in Heart D)

```{r}
all_bottom_MElaSCL100_df_Heart_D<-filter(all_bottom_MElaSCL100_df, organ=="Heart_D")
bottom_metabolite<-all_bottom_MElaSCL100_df_Heart_D$ID 
column_list <- list()
for (name in bottom_metabolite) {
  column_list[[name]] <- metadata[[name]] 
} 

new_df <- as.data.frame(column_list)
new_df<-cbind(new_df, metadata[c(1,2,3, 4, 5)])

new_df <- new_df %>%
  filter(ATTRIBUTE_parasite_species!="CL_Luc", ATTRIBUTE_infected_parasites!="100") %>%
  filter(ATTRIBUTE_organ=="Heart_D") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='1000' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

#boxplot comparing peak areas of each top MElaS metabolites by post infection days and condition. A box plot is produced for each metabolite 
for (metabolite in bottom_metabolite) {
  print(metabolite)
  
  plot<-ggplot(new_df, aes(x = as.factor(ATTRIBUTE_endpoint), y = (.data[[metabolite]]), fill = as.factor(ATTRIBUTE_infected_parasites))) +
  geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
  labs(title=paste0("Peak Area of Bottom MElaS ", metabolite), x = "Endpoint (weeks)", y = "Peak Area", fill="Infection Status") + geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9) +theme_bw() + scale_fill_manual(values = c("infected" = "red", "uninfected" = "grey")) +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(legend.text=element_text(size=13))+  theme(legend.title=element_text(size=rel(1.8)))
  print(plot)
  plotname <- paste0("images2/bottom_metabolites_plots/", metabolite, "boxplot_bottom.png")
  ggsave(plotname, dpi = 300)
}
```

## Plot lineplots of top 10 Metabolites and there FC at FC_0w, FC_3w and LogFC12w

```{r}
all_top_MElaSCL100_df<-filter(all_top_MElaSCL100_df, organ=="Heart_D")
for (organ in unique(all_top_MElaSCL100_df$organ)) {
  temp <- all_top_MElaSCL100_df[all_top_MElaSCL100_df$organ == organ,]
  
  long_noflatData_with_time_top <- pivot_longer(temp, cols = c(FC_0w, FC_3w, FC_12w), 
                          names_to = "Time", values_to = "LogFC")
  long_noflatData_with_time_top$Time <- factor(long_noflatData_with_time_top$Time, levels=c("FC_0w", "FC_3w", "FC_12w"))
  
  for (metabolite in unique(long_noflatData_with_time_top$ID)) {
    temp<-long_noflatData_with_time_top[long_noflatData_with_time_top$ID==metabolite, ]
    p=ggplot(temp, aes(x = Time, y = LogFC, group=1)) + geom_line() + geom_point() + labs(title = paste0("Top MElas: ", metabolite, " FC Between Endpoints in ", organ), x = "Endpoint (weeks)", y = "FC")+ theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(2.8))) + theme(axis.title.x = element_text(size = rel(1.8))) +theme(axis.title.x = element_blank(), axis.text.x = element_blank(), title= element_blank())
  print(p)
  plotname2 <- paste0("images2/top_metabolites_plots/", metabolite, "lineplot.png")
  ggsave(plotname2, dpi = 300) 

  }
}
```

## Plot lineplots of bottom 10 Metabolites and there FC at FC_0w, FC_3w and LogFC12w

```{r}
all_bottom_MElaSCL100_df<-filter(all_bottom_MElaSCL100_df, organ=="Heart_D")
for (organ in unique(all_bottom_MElaSCL100_df$organ)) {
  temp <- all_bottom_MElaSCL100_df[all_bottom_MElaSCL100_df$organ == organ,]
  
  long_noflatData_with_time_bottom <- pivot_longer(temp, cols = c(FC_0w, FC_3w, FC_12w), 
                          names_to = "Time", values_to = "LogFC")
  long_noflatData_with_time_bottom$Time <- factor(long_noflatData_with_time_bottom$Time, levels=c("FC_0w", "FC_3w", "FC_12w"))
  
  for (metabolite in unique(long_noflatData_with_time_bottom$ID)) {
    temp<-long_noflatData_with_time_bottom[long_noflatData_with_time_bottom$ID==metabolite, ]
    p=ggplot(temp, aes(x = Time, y = LogFC, group=1)) + geom_line() + geom_point() + labs(title = paste0("Bottom MElas: ", metabolite, " FC Between Endpoints in ", organ), x = "Endpoint Rations", y = "FC")+ theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(2.8))) + theme(axis.title.x = element_text(size = rel(1.8))) +theme(axis.title.x = element_blank(), axis.text.x = element_blank(), title= element_blank())
  print(p)
  plotname2 <- paste0("images2/", metabolite, "lineplot_bottom.png")
  ggsave(plotname2, dpi = 300)
  }
}

```

# Impact of inoculum  and T. cruzi strain on MElaS values
## Run a Benjamini–Hochberg FDR corrected Wilcoxon signed-rank test comparing by parasite strains

```{r}
noflatData<- noflatData %>%
  mutate(log_MElaS = log10(MElaS)) %>%
  filter(MElaS !=0)

noflatData$Group <- paste0(noflatData$parasite_species, "_", noflatData$infected_parasites)

p_values_df <- data.frame(Organ = character(),
                          Group1 = character(),
                          Group2 = character(),
                          Test = character(),
                          P_Value = numeric(),
                          Fold_Change = numeric(),
                          stringsAsFactors = FALSE)

OrganList <- c(unique(noflatData$organ))
for (Organ in OrganList){
  print(Organ)
  temp <- noflatData[noflatData$organ == Organ,]
  comparisons <- combn(levels(factor(temp$Group)), 2, simplify = FALSE)
  comparisons <- c(comparisons[2], comparisons[5])
    for (comp in comparisons) {
      group1 <- comp[1]
      group2 <- comp[2]
      
      # Subset data for the two strains
      temp_subset <- temp[temp$Group %in% c(group1, group2), ]
      
      if (length(unique(temp_subset[temp_subset$Group == group1,]$log_MElaS)) > 1 & length(temp_subset[temp_subset$Group == group1,]$log_MElaS) > 2) {
        normalcy1 <- shapiro.test(temp_subset[temp_subset$Group == group1,]$log_MElaS)$p.value
        } else {
          normalcy1 <- 0 
        }
      if (length(unique(temp_subset[temp_subset$Group == group2,]$log_MElaS)) > 1 & length(temp_subset[temp_subset$Group == group2,]$log_MElaS) > 2) {
        normalcy2 <- shapiro.test(temp_subset[temp_subset$Group == group2,]$log_MElaS)$p.value
        } else {
          normalcy2 <- 0 
        }
    
      # Choose the appropriate test based on normality
      if (normalcy1 < 0.05 | normalcy2 < 0.05) {
        test_used <- "Wilcoxon"
        plotTest <- "wilcox.test"
        p_val <- wilcox.test(log_MElaS ~ Group, data = temp_subset)$p.value
      } else {
        test_used <- "T-test"
        plotTest <- "t.test"
        p_val <- t.test(log_MElaS ~ Group, data = temp_subset)$p.value
      }
      
      #make fold change
      fc <- median(temp_subset[temp_subset$Group == group1,]$log_MElaS)/median(temp_subset[temp_subset$Group == group2,]$log_MElaS)
      
      # Store results in the data frame
      p_values_df <- rbind(p_values_df, data.frame(
        Organ = Organ,
        Group1 = group1,
        Group2 = group2,
        Test = test_used,
        Fold_Change = fc,
        P_Value = p_val
      ))
    }
}
#p-value adjustment
p_values_df$p_value_adj <- 
  p.adjust(p_values_df$P_Value, method = "BH")

#write.csv(p_values_df, "by_parasite_group_pvalues.csv", row.names = FALSE)
```

## Run a Benjamini–Hochberg FDR corrected Wilcoxon signed-rank test comparing by inoculum

```{r}
p_values_df_2 <- data.frame(Organ = character(),
                          Group1 = character(),
                          Group2 = character(),
                          Test = character(),
                          Fold_Change = numeric(),
                          P_Value = numeric(),
                          stringsAsFactors = FALSE)

OrganList <- c(unique(noflatData$organ))
for (Organ in OrganList){
  print(Organ)
  temp <- noflatData[noflatData$organ == Organ,]
  comparisons <- combn(levels(factor(temp$Group)), 2, simplify = FALSE)
  comparisons <- c(comparisons[1], comparisons[6])
    for (comp in comparisons) {
      group1 <- comp[1]
      group2 <- comp[2]
      
      # Subset data for the two inoculums
      temp_subset <- temp[temp$Group %in% c(group1, group2), ]
      
      if (length(unique(temp_subset[temp_subset$Group == group1,]$log_MElaS)) > 1 & length(temp_subset[temp_subset$Group == group1,]$log_MElaS) > 2) {
        normalcy1 <- shapiro.test(temp_subset[temp_subset$Group == group1,]$log_MElaS)$p.value
        } else {
          normalcy1 <- 0
        }
      if (length(unique(temp_subset[temp_subset$Group == group2,]$log_MElaS)) > 1 & length(temp_subset[temp_subset$Group == group2,]$log_MElaS) > 2) {
        normalcy2 <- shapiro.test(temp_subset[temp_subset$Group == group2,]$log_MElaS)$p.value
        } else {
          normalcy2 <- 0 
        }
    
      # Choose the appropriate test based on normality
      if (normalcy1 < 0.05 | normalcy2 < 0.05) {
        test_used <- "Wilcoxon"
        plotTest <- "wilcox.test"
        p_val <- wilcox.test(log_MElaS ~ Group, data = temp_subset, p.adjust.method = "BH")$p.value
      } else {
        test_used <- "T-test"
        plotTest <- "t.test"
        p_val <- t.test(log_MElaS ~ Group, data = temp_subset)$p.value
      }
      
      # make FC
      fc <- median(temp_subset[temp_subset$Group == group1,]$log_MElaS)/median(temp_subset[temp_subset$Group == group2,]$log_MElaS)
      
      # Store results in the data frame
      p_values_df_2 <- rbind(p_values_df_2, data.frame(
        Organ = Organ,
        Group1 = group1,
        Group2 = group2,
        Test = test_used,
        Fold_Change = fc,
        P_Value = p_val
      ))
    }
    
}
#p-value adjustment
p_values_df_2$p_value_adj <- 
  p.adjust(p_values_df_2$P_Value, method = "BH")

#write.csv(p_values_df_2, "by_parasite_conc_pvalues.csv", row.names = FALSE)
```

## Visulatization boxplots log-scaled MElaS values

```{r}
#Figure 7A. Log-scaled MElaS values for each organ by inoculation concentration 100
# Colors
fill_colors <- c(
  "CL"  = "lightblue", 
  "JR"  = "salmon"
)

noflatData$strain_conc <- paste0(
  noflatData$parasite_species, "_",
  noflatData$infected_parasites
)
noflatData_100 <- noflatData %>%
  mutate(log_MElaS = log10(MElaS)) %>%
  filter(infected_parasites=="100")

noflatData_100$organ <- factor(noflatData_100$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(data=noflatData_100, aes(x = as.factor(organ), y = log_MElaS, fill = parasite_species)) + geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) + stat_compare_means(
    aes(group = parasite_species),
    method = "wilcox.test",
    label = "p.signif",
    hide.ns = TRUE,
    p.adjust.method = "BH",
    bracket.size = 0.8,
    label.y = 1.8
  )  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ labs(x="Organ", y=expression(Log [10]~MElaS))+
  
  scale_fill_manual(values = fill_colors,
                    name = "Strain") + 
  coord_cartesian(ylim = c(-3.5, 2)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.8)),
    axis.title.y = element_text(size = rel(1.8)),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = rel(1.3)),
    plot.title = element_text(size = 20)
  )
#Figure 7B. Log-scaled MElaS values for each organ by inoculation concentration 1000
# Colors
fill_colors2 <- c(
  "CL"  = "steel blue", 
  "JR"  = "red3"
)
noflatData_1000 <- noflatData %>%
  mutate(log_MElaS = log10(MElaS)) %>%
  filter(infected_parasites=="1000")

noflatData_1000$organ <- factor(noflatData_1000$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(data=noflatData_1000, aes(x = as.factor(organ), y = log_MElaS, fill = parasite_species)) + geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) + stat_compare_means(
    aes(group = parasite_species),
    method = "wilcox.test",
    label = "p.signif",
    hide.ns = TRUE,
    p.adjust.method = "BH",
    bracket.size = 0.8,
    label.y = 1.8
  )  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ labs(x="Organ", y=expression(Log [10]~MElaS))+
  
  scale_fill_manual(values = fill_colors2,
                    name = "Strain") + 
  coord_cartesian(ylim = c(-3.5, 2)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.8)),
    axis.title.y = element_text(size = rel(1.8)),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = rel(1.3)),
    plot.title = element_text(size = 20)
  )

#Figure 8A. Log-scaled MElaS values for each organ by parasite strain CL
# Colors
fill_colors3 <- c(
  "100"  = "lightblue", 
  "1000"  = "steel blue"
)


noflatData_CL <- noflatData %>%
  mutate(log_MElaS = log10(MElaS)) %>%
  filter(parasite_species=="CL")

noflatData_CL$organ <- factor(noflatData_CL$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(data=noflatData_CL, aes(x = as.factor(organ), y = log_MElaS, fill = infected_parasites)) + geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) + stat_compare_means(
    aes(group = infected_parasites),
    method = "wilcox.test",
    label = "p.signif",
    hide.ns = TRUE,
    p.adjust.method = "BH",
    bracket.size = 0.8,
    label.y = 1.8
  )  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ labs(x="Organ", y=expression(Log [10]~MElaS))+
  
  scale_fill_manual(values = fill_colors3,
                    name = "Concentration") + 
  coord_cartesian(ylim = c(-3.5, 2)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.8)),
    axis.title.y = element_text(size = rel(1.8)),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = rel(1.3)),
    plot.title = element_text(size = 20)
  )
#Figure 8B. Log-scaled MElaS values for each organ by parasite strain JR
# Colors
fill_colors4 <- c(
  "100"  = "salmon", 
  "1000"  = "red3"
)

noflatData_JR <- noflatData %>%
  mutate(log_MElaS = log10(MElaS)) %>%
  filter(parasite_species=="JR")

noflatData_JR$organ <- factor(noflatData_JR$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(data=noflatData_JR, aes(x = as.factor(organ), y = log_MElaS, fill = infected_parasites)) + geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) + stat_compare_means(
    aes(group = infected_parasites),
    method = "wilcox.test",
    label = "p.signif",
    hide.ns = TRUE,
    p.adjust.method = "BH",
    bracket.size = 0.8,
    label.y = 1.8
  )  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ labs(x="Organ", y=expression(Log [10]~MElaS))+
  
  scale_fill_manual(values = fill_colors4,
                    name = "Concentration") + 
  coord_cartesian(ylim = c(-3.5, 2)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.8)),
    axis.title.y = element_text(size = rel(1.8)),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = rel(1.3)),
    plot.title = element_text(size = 20)
  )
```


# Impact of inoculation concentration and T. cruzi strain on the proportion of elastic metabolites

```{r}
#total percentage of nonzeros
non_zero_count=sum(noflatData$MElaS !=0)
total_rows=nrow(noflatData)
proportion_non_zero <- (non_zero_count / total_rows) *100
zero_count=total_rows-non_zero_count

#percentage of nonzeros by organ 
proportion_non_zero_by_organ <- dataWithMElaSTypes %>%
  group_by(organ, parasite_species, infected_parasites) %>%
  summarise(
    non_zero_count_organ=sum(MElaS !=0),
    total_rows_organ=n(),
    proportion_non_zero_organ = (non_zero_count_organ / total_rows_organ) *100,
    proportion_non_zero_organ =format(round(proportion_non_zero_organ, 3), nsmall = 3),
    zero_count=total_rows_organ-non_zero_count_organ
  )

proportion_non_zero_by_organ$proportion_non_zero_organ=as.numeric(proportion_non_zero_by_organ$proportion_non_zero_organ)
```

## Percentage of Non-Zero MElas Scores per Organ Barplots

```{r}

proportion_non_zero_by_organ$strain_conc <- paste0(
  proportion_non_zero_by_organ$parasite_species, "_",
  proportion_non_zero_by_organ$infected_parasites
)
proportion_non_zero_by_organ$organ <- factor(proportion_non_zero_by_organ$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

#Figure 5A. Comparison of the proportion of metabolites with a MElaS value of nonzero for each organ by parasite strain
# Colors
fill_colors <- c(
  "CL_100"  = "lightblue", 
  "CL_1000" = "steel blue",
  "JR_100"  = "salmon",
  "JR_1000" = "red3"
)

ggplot(data = proportion_non_zero_by_organ,
       aes(x = as.factor(organ),
           y = proportion_non_zero_organ,
           fill = strain_conc)) +
  geom_col(position = position_dodge(width = 0.9)) + facet_wrap(vars(infected_parasites)) +
  
  scale_fill_manual(values = fill_colors,
                    name = "Strain and \nConcentration") +
  
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.8)),
    axis.title.y = element_text(size = rel(1.8)),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = rel(1.3))
  ) +
  labs(x = "Organ", y = "Percentage (%)")

#Figure 5B. Comparison of the proportion of metabolites with a MElaS value of nonzero for each organ by inoculation concentration
# Colors
fill_colors <- c(
  "CL_100"  = "lightblue", 
  "CL_1000" = "steel blue",
  "JR_100"  = "salmon",
  "JR_1000" = "red3"
)

ggplot(data = proportion_non_zero_by_organ,
       aes(x = as.factor(organ),
           y = proportion_non_zero_organ,
           fill = strain_conc)) +
  geom_col(position = position_dodge(width = 0.9)) + facet_wrap(vars(parasite_species)) +
  
  scale_fill_manual(values = fill_colors,
                    name = "Strain and \nConcentration") +
  
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.8)),
    axis.title.y = element_text(size = rel(1.8)),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = rel(1.3))
  ) +
  labs(x = "Organ", y = "Percentage (%)")


```

## Fisher's Test: comparing JR and CL at 100 and 1000 of each organ 

```{r}
fisher_test_results_byJRvCL <- data.frame(
  Organ = character(),
  P_Value = numeric(),
  Load = character(),
  Parasite = character(),
  FoldChange = numeric(),
  stringsAsFactors = FALSE
)

for (Tissue in unique(proportion_non_zero_by_organ$organ)) {
  proportion_non_zero_by_Organ=proportion_non_zero_by_organ %>%
  group_by(parasite_species, infected_parasites, organ) %>%
  filter(organ==Tissue)
  
  for (load in unique(proportion_non_zero_by_Organ$infected_parasites)) {
    noflatDataframe_infected_parasites=proportion_non_zero_by_Organ %>%
      filter(infected_parasites==load) %>%
      select(non_zero_count_organ, zero_count, proportion_non_zero_organ)
    print(noflatDataframe_infected_parasites)
    foldchange = noflatDataframe_infected_parasites$proportion_non_zero_organ[1] / noflatDataframe_infected_parasites$proportion_non_zero_organ[2]
    
    table_counts <- noflatDataframe_infected_parasites %>%
      group_by(parasite_species, infected_parasites) %>%
      summarise(non_zero = sum(non_zero_count_organ),
                zero = sum(zero_count)) %>%
      ungroup()
    
    contingency <- as.matrix(table_counts[, c("non_zero", "zero")])
    rownames(contingency) <- table_counts$parasite_species
    print(contingency)
    
    fisher_result <- fisher.test(contingency, workspace=2000000000)
    print(paste0("Fisher Test Comparing JR and CL at ", " Parasite Load in the ", Tissue))
    print(fisher_result)
    
    fisher_test_results_byJRvCL <- rbind(
      fisher_test_results_byJRvCL,
      data.frame(Load = load, p_value = fisher_result$p.value, Organ = Tissue, parasite=paste(rownames(contingency), collapse="_vs_"), FoldChange=foldchange)
    )

  }
}
#p-value adjusted
fisher_test_results_byJRvCL$p_value_adj <- 
  p.adjust(fisher_test_results_byJRvCL$p_value, method = "BH")

# Save full results
write.csv(fisher_test_results_byJRvCL, 
          "fisher_test_results_byJRvCL.csv", 
          row.names = FALSE)
```

## Fisher's Test: comparing 100 and 1000 at JR and CL of each organ

```{r}
fisher_test_results_by100v1000 <- data.frame(
  Organ = character(),
  P_Value = numeric(),
  Load = character(),
  parasite = character(),
  FoldChange = numeric(),
  stringsAsFactors = FALSE
)

for (Tissue in unique(proportion_non_zero_by_organ$organ)) {
  proportion_non_zero_by_Organ_100v1000=proportion_non_zero_by_organ %>%
  filter(organ==Tissue) %>%  
  group_by(parasite_species, infected_parasites, organ)
  
  for (Parasite in unique(proportion_non_zero_by_organ$parasite_species)) {
    dataframe_Parasite=proportion_non_zero_by_Organ_100v1000 %>%
      filter(parasite_species==Parasite) %>%
      select(non_zero_count_organ, zero_count, infected_parasites, proportion_non_zero_organ)
    #print(dataframe_Parasite)
    foldchange = dataframe_Parasite$proportion_non_zero_organ[1] / dataframe_Parasite$proportion_non_zero_organ[2]
    
    table_counts <- dataframe_Parasite %>%
      group_by(infected_parasites) %>%
      summarise(non_zero = sum(non_zero_count_organ),
                zero = sum(zero_count)) %>%
      ungroup()
    
    contingency <- as.matrix(table_counts[, c("non_zero", "zero")])
    rownames(contingency) <- table_counts$infected_parasites
    print(contingency)
    
    fisher_result <- fisher.test(contingency, workspace=2000000000)
    print(paste0("Fisher Test Comparing 100 and 1000 Parasite Load with in the ", Tissue))
    print(fisher_result)
    
    fisher_test_results_by100v1000 <- rbind(
      fisher_test_results_by100v1000,
      data.frame(Load = paste(rownames(contingency), collapse="_vs_"), P_value = fisher_result$p.value, Organ = Tissue, parasite_species=Parasite, FoldChange=foldchange)
    )

  }
}
fisher_test_results_by100v1000$p_value_adj <- 
  p.adjust(fisher_test_results_by100v1000$P_value, method = "BH")

#Save full results
write.csv(fisher_test_results_by100v1000, 
          "fisher_test_results_by100v1000.csv", 
          row.names = FALSE)
```

# Differences in non-zero MElaS values by sampling site

```{r}
#JR 1000

#Kruskal Wallis test to compare significance among organs
kruskalwalisJR1000 <- kruskal.test(MElaS ~ organ, data = JR_1000_noflatData)
kruskal.test(MElaS ~ organ, data = JR_1000_noflatData) #organs are significantly different p<2.2e-16

 #Benjamini–Hochberg FDR corrected Dunn’s post hoc multiple comparison test
dunntestJR1000 <- dunnTest(MElaS ~ organ,
         data=JR_1000_noflatData,
         method="bh")
dunn_df_JR1000 <- dunntestJR1000$res

JR_1000_noflatData$organ <- factor(JR_1000_noflatData$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

#boxplots by organ
boxplot_allorgans_JR1000 <- ggplot(JR_1000_noflatData, aes(x=organ, y=MElaS, fill=organ))+ geom_boxplot() +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) +  labs(x="Organ", y="MElaS") + theme(legend.position="none")
print(boxplot_allorgans_JR1000)
#plotname <- "thesis_pres_images/defense_images/boxplot_allorgans_JR1000.png"
#ggsave(plotname, dpi = 300)


#JR 100
#Kruskal Wallis test to compare significance among organs
kruskal.test(MElaS ~ organ, data = JR_100_noflatData) 

 #Benjamini–Hochberg FDR corrected Dunn’s post hoc multiple comparison test
dunntestJR100 <- dunnTest(MElaS ~ organ,
         data=JR_100_noflatData,
         method="bh")
dunn_df_JR100 <- dunntestJR100$res

JR_100_noflatData$organ <- factor(JR_100_noflatData$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

#boxplots by organ
boxplot_allorgans_JR100 <- ggplot(JR_100_noflatData, aes(x=organ, y=MElaS, fill=organ))+ geom_boxplot() +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(plot.title=element_text(size=25)) +  labs(x="Organ", y="MElaS") + theme(legend.position="none")
print(boxplot_allorgans_JR100)
#plotname <- "thesis_pres_images/defense_images/boxplot_allorgans_JR100.png"
#ggsave(plotname, dpi = 300)

#CL 1000
#Kruskal Wallis test to compare significance among organs
kruskal.test(MElaS ~ organ, data = CL_1000_noflatData) 

 #Benjamini–Hochberg FDR corrected Dunn’s post hoc multiple comparison test
dunntestCL1000 <- dunnTest(MElaS ~ organ,
         data=CL_1000_noflatData,
         method="bh")
dunn_df_CL1000 <- dunntestCL1000$res

CL_1000_noflatData$organ <- factor(CL_1000_noflatData$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

#boxplots by organ
boxplot_allorgans_CL1000 <- ggplot(CL_1000_noflatData, aes(x=organ, y=MElaS, fill=organ))+ geom_boxplot() +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(plot.title=element_text(size=25)) +  labs(x="Organ", y="MElaS") + theme(legend.position="none")
print(boxplot_allorgans_CL1000)
#plotname <- "thesis_pres_images/defense_images/boxplot_allorgans_CL1000.png"
#ggsave(plotname, dpi = 300)


#CL 100
#Kruskal Wallis test to compare significance among organs
kruskal.test(MElaS ~ organ, data = CL_100_noflatData) 

 #Benjamini–Hochberg FDR corrected Dunn’s post hoc multiple comparison test
dunntestCL100 <-dunnTest(MElaS ~ organ,
         data=CL_100_noflatData,
         method="bh", list = TRUE)
dunn_df_CL100 <- dunntestCL100$res

CL_100_noflatData$organ <- factor(CL_100_noflatData$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

#boxplots by organ
boxplot_allorgans_CL100 <- ggplot(CL_100_noflatData, aes(x=organ, y=MElaS, fill=organ))+ geom_boxplot() +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(plot.title=element_text(size=25)) +  labs(x="Organ", y="MElaS") + theme(legend.position="none")
print(boxplot_allorgans_CL100)
#plotname <- "thesis_pres_images/defense_images/boxplot_allorgans_CL100.png"
#ggsave(plotname, dpi = 300)

#Save full results
write.csv(dunn_df_JR1000, "dunn_df_JR1000.csv", row.names = FALSE)
write.csv(dunn_df_JR100, "dunn_df_JR100.csv", row.names = FALSE)
write.csv(dunn_df_CL1000, "dunn_df_CL1000.csv", row.names = FALSE)
write.csv(dunn_df_CL100, "dunn_df_CL100.csv", row.names = FALSE)

```

#comparing types (not done yet)

```{r}
# do i need dont have dataWithMElaSTypesCountnoflat
#counts of type by parasite and concentration 
for (organ in unique(dataWithMElaSTypesCountnoflat$organ)) {
  temp <- dataWithMElaSTypesCountnoflat[dataWithMElaSTypesCountnoflat$organ == organ,]
  plot<-ggplot(data=temp, aes(x = as.factor(Type), y = TypeCount, fill = as.factor(parasite_species))) +
  geom_col(position = "dodge") +
  facet_wrap(vars(infected_parasites)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle(paste0("Count of Categories in ", organ)) + ylim(0, 700) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.8)),
    axis.title.y = element_text(size = rel(1.8)),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = rel(1.8)),
    plot.title = element_text(size = 20)
  ) +
  labs(x = "Organ", y = "Count", fill="Parasite Strain")
  print(plot)
}

dataWithMElaSTypesCountnoflat$strain_conc <- paste0(
  dataWithMElaSTypesCountnoflat$parasite_species, "_",
  dataWithMElaSTypesCountnoflat$infected_parasites
)


# Colors
fill_colors <- c(
  "CL_100"  = "lightblue", 
  "CL_1000" = "blue3",
  "JR_100"  = "firebrick1",
  "JR_1000" = "red3"
)

# do i need dont have dataWithMElaSTypesCountnoflat
for (organ in unique(dataWithMElaSTypesCountnoflat$organ)) {
  temp <- dataWithMElaSTypesCountnoflat[dataWithMElaSTypesCountnoflat$organ == organ,]
  plot<-ggplot(data=temp, aes(x = as.factor(Type), y = TypeCount, fill = strain_conc)) +
  geom_col(position = "dodge") +
  facet_wrap(vars(infected_parasites)) + scale_fill_manual(values = fill_colors,
                    name = "Strain and Load")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle(paste0("Count of Categories in ", organ)) + ylim(0, 700) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.8)),
    axis.title.y = element_text(size = rel(1.8)),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = rel(1.8)),
    plot.title = element_text(size = 20)
  ) +
  labs(x = "Organ", y = "Count")
  print(plot)
}

```
