---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
```{r}
library(dplyr) # version 1.1.4
library(ggplot2) # version 3.5.2
library(tidyr) # version 1.3.1
library(ggpubr) # version 0.6.0
library(FSA)
library(dunn.test)
library(purrr)
library(stringr)
library(viridis) 
```
#load data
```{r}
data<-read.csv("ALL_DATA.csv") #can change back to ALL_DATA.csv for oringial tranformation
data$infected_parasites=as.factor(data$infected_parasites)

```
#Filter Data

```{r}
#works 10/15??
dataWithMElaSTypes <- data %>%
  mutate(
    Type = case_when(
      (logFC_3v0 >= -0.3 & logFC_3v0 <= 0.3) & (logFC_12v3>=-0.3 & logFC_12v3 <= 0.3)  ~ "flat",
      (logFC_3v0 >= -0.3 & logFC_3v0 <= 0.3) & (logFC_12v3 > 0.3) ~ "flat_up",
      (logFC_3v0 >= -0.3 & logFC_3v0 <= 0.3) & (logFC_12v3 < -0.3) ~ "flat_down",
      (logFC_3v0 > 0.3) & (logFC_12v3>=-0.3 & logFC_12v3<=0.3) ~ "up_flat",
      (logFC_3v0 < -0.3) & (logFC_12v3>=-0.3 & logFC_12v3<=0.3) ~ "down_flat",
      (logFC_12v3 > 0.3 & logFC_3v0 > 0.3) ~ "up_up",
      (logFC_12v3 < -0.3 & logFC_3v0 < -0.3) ~ "down_down",
      (logFC_3v0 > 0.3) & (logFC_12v3 < -0.3) ~ "up_down",
      (logFC_3v0 < -0.3) & (logFC_12v3 > 0.3) ~ "down_up"
    )
  )
# for 24 weeks analysis
dataWithMElaSTypes <- dataWithMElaSTypes %>%
  mutate(
    Type_24w = case_when(FC_24w >= FC_12w ~ "up", FC_24w <= FC_12w ~ "down"))
      
noflatData<-filter(dataWithMElaSTypes, Type!="flat", )

dataWithMElaSTypes$organ <- factor(dataWithMElaSTypes$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))
#write.csv(dataWithMElaSTypes, "dataWithMElaSTypes.csv")

dataWithMElaSTypes <- dataWithMElaSTypes %>%
  mutate(extracted_ID = str_remove_all(ID, "X(\\d+).(\\d+)_(\\d+).(\\d+)_"))
```

```{r}
#overlap of metabolites among different organs
noflatData_overlapping <- noflatData %>%
  group_by(ID) %>%
  mutate(count_occurrences = n()) %>%
  ungroup() 


# most overlapped top metabolite

all_top_df <- data.frame()
for (organ in unique(dataWithMElaSTypes$organ)) {
  temp <- dataWithMElaSTypes[dataWithMElaSTypes$organ == organ,]
  
  top<-top_n(temp, 50, MElaS) #metabolites with top MElaS scores
  all_top_df <- rbind(all_top_df, top)
}

top_overlap <- all_top_df %>%
  group_by(ID) %>%
  mutate(count_occurrences = n()) %>%
  ungroup()


all_bottom_df <- data.frame()
for (organ in unique(dataWithMElaSTypes$organ)) {
  temp <- dataWithMElaSTypes[dataWithMElaSTypes$organ == organ,]
  
  bottom<-top_n(temp, -50, MElaS) #metabolites with lowest MElaS scores
  bottom <- bottom[sample(nrow(bottom), 50), ]
  all_bottom_df <- rbind(all_bottom_df, bottom)
}
```

```{r}
JR_1000_noflatData<- noflatData %>%
  filter(parasite_species=="JR", infected_parasites=="1000")

JR_100_noflatData<- noflatData %>%
  filter(parasite_species=="JR", infected_parasites=="100")

CL_1000_noflatData<- noflatData %>%
  filter(parasite_species=="CL", infected_parasites=="1000")

CL_100_noflatData<- noflatData %>%
  filter(parasite_species=="CL", infected_parasites=="100")
```

#JR, 1000 Analysis

```{r}
#Top MElaS
noflatDatalist = list()
for (organ in unique(JR_1000_noflatData$organ)) {
  temp <- JR_1000_noflatData[JR_1000_noflatData$organ == organ,]
  
  top<-top_n(temp, 10, MElaS) #metabolites with top MElaS scores
  
  long_noflatData <- pivot_longer(top, cols = c(logFC_3v0, logFC_12v3),names_to = "Ratio", values_to = "LogFC")
  long_noflatData$Ratio <- gsub("Ratio", "", long_noflatData$Ratio)
  p=ggplot(long_noflatData, aes(x = Ratio, y = LogFC, group = ID, color=ID)) + geom_line(alpha = 0.6) + geom_point() + labs(title = paste0("Top Ten MElas Endpoint Ratios in ", organ), x = "Endpoint Ratio", y = "lnFC", color="Metabolites", group="Metabolites")+ theme_minimal() 
  p +  theme(axis.text=element_text(size=10), axis.title=element_text(size=20),plot.title = element_text(size=22, hjust=0.5))+theme(legend.text = element_text(size = 10), legend.title = element_text(size = 15))
  print(p)
  noflatDatalist[[length(noflatDatalist) + 1]] <- top
}
all_top_MElaSJR1000_df<-do.call(rbind, noflatDatalist)
```

```{r}
#Bottom MElaS
noflatDatalist2 = list()
for (organ in unique(JR_1000_noflatData$organ)) {
  temp <- JR_1000_noflatData[JR_1000_noflatData$organ == organ,]
  
  bottom<-top_n(temp, -10, MElaS) #metabolites with lowest MElaS scores
  bottom <- bottom[sample(nrow(bottom), 10), ] #randomly sample 10
  long_noflatData <- pivot_longer(bottom, cols = c(logFC_3v0, logFC_12v3), 
                          names_to = "Ratio", values_to = "LogFC")
  long_noflatData$Ratio <- gsub("Ratio", "", long_noflatData$Ratio)
  p=ggplot(long_noflatData, aes(x = Ratio, y = LogFC, group = ID, color=ID)) +
  geom_line(alpha = 0.6) +
  geom_point() +
  labs(title = paste0("Bottom Ten MElas EndpointRatios in ", organ),
       x = "Endpoint Ratio", y = "lnFC", group="Metabolite", color="Metabolite") +
  theme_minimal()
  p+  
		theme(axis.text=element_text(size=10),
				axis.title=element_text(size=20),plot.title = element_text(size=22, hjust=0.5))+theme(legend.text = element_text(size = 10), legend.title = element_text(size = 15))
  print(p)
  noflatDatalist2[[length(noflatDatalist2) + 1]] <- bottom

}
all_bottom_MElaSJR1000_df<-do.call(rbind, noflatDatalist2)
```
#boxplot comparing peak areas of each top MElaS metabolites by post infection days and condition
```{r}
#load metadata
metadata<-read.csv("DisTol_mergeddata.csv") #load data can use CLR_transformed_metadata or DisTol_mergeddata
metadata<-na.omit(metadata)
ATTRIBUTE_infected_parasites<-as.factor(metadata$ATTRIBUTE_infected_parasites)
ATTRIBUTE_endpoint<-as.factor(metadata$ATTRIBUTE_endpoint)
#metadata<-filter(metadata, ATTRIBUTE_endpoint!="24w", )
metadata<-filter(metadata, ATTRIBUTE_infected_parasites!='10000', )

#only if you are doing CLR transformation
#my_data <- metadata[, -c(1, 2, 3, 4, 5, 6)]
#min(my_data[my_data != 0]) #just gives you the minimum


#data_rclr_min_value <- min(my_data)
#my_data[my_data == 0] = data_rclr_min_value #change 0w values to min instead of 0

#metadata <- merge(
 # metadata[, 1:5],
  #my_data,
 # by = "row.names"
#)

#rownames(metadata) <- metadata$Row.names
#metadata$Row.names <- NULL
#which(names(metadata)=="bar")
```

```{r}
JR_1000_metadata <- metadata %>%
  filter(ATTRIBUTE_parasite_species!="CL_Luc", ATTRIBUTE_infected_parasites!="100") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='1000' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w", "24w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

```

#code for visualization of top metabolite boxplots (for presentation)

```{r}
all_top_MElaSJR1000_df_Heart_D<-filter(all_top_MElaSJR1000_df, organ=="Heart_D")
top_metabolite<-all_top_MElaSJR1000_df_Heart_D$ID 
column_list <- list()
for (name in top_metabolite) {
  column_list[[name]] <- metadata[[name]] 
} #make a dataset that only includes metabolites with top MElas scores

new_df <- as.data.frame(column_list)
new_df<-cbind(new_df, metadata[c(1,2,3, 4, 5)]) #make new dataset with data set of top 10 metabolites and all samples

new_df <- new_df %>%
  filter(ATTRIBUTE_parasite_species!="CL_Luc", ATTRIBUTE_infected_parasites!="100") %>%
  filter(ATTRIBUTE_organ=="Heart_D") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='1000' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w", "24w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

#boxplot comparing peak areas of each top MElaS metabolites by post infection days and condition. A box plot is produced for each metabolite 
for (metabolite in top_metabolite) {
  print(metabolite)
                                                              #this was changed to log(.data[[metabolite]]+1 dont know why)
  plot<-ggplot(new_df, aes(x = as.factor(ATTRIBUTE_endpoint), y = (.data[[metabolite]]), fill = as.factor(ATTRIBUTE_infected_parasites))) +
  geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
  labs(title=paste0("Peak Area of Top MElaS ", metabolite), x = "Endpoint (weeks)", y = "Peak Area", fill="Infection Status") + geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9) +theme_bw() + scale_fill_manual(values = c("infected" = "red", "uninfected" = "grey")) +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(legend.text=element_text(size=13))+  theme(legend.title=element_text(size=rel(1.8)))
  print(plot)
  plotname <- paste0("images2/", metabolite, "boxplot.png")
  ggsave(plotname, dpi = 300)
  
}
```

#code for visualization of bottom metabolite boxplots (for presentation)

```{r}
all_bottom_MElaSJR1000_df_Heart_D<-filter(all_bottom_MElaSJR1000_df, organ=="Heart_D")
bottom_metabolite<-all_bottom_MElaSJR1000_df_Heart_D$ID 
column_list <- list()
for (name in bottom_metabolite) {
  column_list[[name]] <- metadata[[name]] 
} #make a dataset that only includes metabolites with top MElas scores

new_df <- as.data.frame(column_list)
new_df<-cbind(new_df, metadata[c(1,2,3, 4, 5)]) #make new dataset with data set of top 10 metabolites and all samples

new_df <- new_df %>%
  filter(ATTRIBUTE_parasite_species!="CL_Luc", ATTRIBUTE_infected_parasites!="100") %>%
  filter(ATTRIBUTE_organ=="Heart_D") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='1000' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w", "24w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

#boxplot comparing peak areas of each top MElaS metabolites by post infection days and condition. A box plot is produced for each metabolite 
for (metabolite in bottom_metabolite) {
  print(metabolite)
  
  plot<-ggplot(new_df, aes(x = as.factor(ATTRIBUTE_endpoint), y = (.data[[metabolite]]), fill = as.factor(ATTRIBUTE_infected_parasites))) +
  geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
  labs(title=paste0("Peak Area of Bottom MElaS ", metabolite), x = "Endpoint (weeks)", y = "Peak Area", fill="Infection Status") + geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9) +theme_bw() + scale_fill_manual(values = c("infected" = "red", "uninfected" = "grey")) +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(legend.text=element_text(size=13))+  theme(legend.title=element_text(size=rel(1.8)))
  print(plot)
  plotname <- paste0("images2/", metabolite, "boxplot_bottom.png")
  ggsave(plotname, dpi = 300)
}
```

#boxplot comparing peak areas of each top MElaS metabolites by post infection days and condition
#shows all top metabolites on a plot
```{r}

plots <- all_top_MElaSJR1000_df %>%
  split(.$organ) %>%
  imap(function(df, organ_name) {
    # Get the top metabolite names for this organ
    top_metabolites <- df$ID
    
    # Subset peak data for this organ and those metabolites
    subset_data <- JR_1000_metadata  %>%
      filter(ATTRIBUTE_organ == organ_name) %>%
      select(ATTRIBUTE_infected_parasites, ATTRIBUTE_organ, ATTRIBUTE_endpoint, all_of(top_metabolites)) %>%
      pivot_longer(cols = all_of(top_metabolites),
                   names_to = "metabolite", values_to = "peak_area")
    
    # Make boxplots comparing endpointspoints
    p <- ggplot(subset_data, aes(x = ATTRIBUTE_endpoint, y = peak_area, fill = ATTRIBUTE_infected_parasites)) +
      geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
      facet_wrap(~ metabolite, scales = "free_y", ncol = 5) +
      theme_bw(base_size = 9) +
      labs(title = paste("Top 10 Metabolites in", organ_name, 'for JR, 1000'),
           y = "Peak Area", x = "Endpoint", fill="Infection Status") +
      theme(legend.position = "top",
            strip.text = element_text(face = "bold"))+ geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9) 
    
    print(p)
  })
```

#boxplot comparing peak areas of each bottom MElaS metabolites by post infection days and condition
#shows all bottom metabolites on a plot

```{r}
plots <- all_bottom_MElaSJR1000_df %>%
  split(.$organ) %>%
  imap(function(df, organ_name) {
    # Get the top metabolite names for this organ
    bottom_metabolites <- df$ID
    
    # Subset peak data for this organ and those metabolites
    subset_data <- JR_1000_metadata  %>%
      filter(ATTRIBUTE_organ == organ_name) %>%
      select(ATTRIBUTE_infected_parasites, ATTRIBUTE_organ, ATTRIBUTE_endpoint, all_of(bottom_metabolites)) %>%
      pivot_longer(cols = all_of(bottom_metabolites),
                   names_to = "metabolite", values_to = "peak_area")
    
    # Make boxplots comparing timepoints
    p <- ggplot(subset_data, aes(x = ATTRIBUTE_endpoint, y = peak_area, fill = ATTRIBUTE_infected_parasites)) +
      geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
      facet_wrap(~ metabolite, scales = "free_y", ncol = 5) +
      theme_bw(base_size = 9) +
      labs(title = paste("Bottom 10 Metabolites in", organ_name, 'for JR, 1000'),
           y = "Peak Area", x = "Timepoint", fill="Infection Status") +
      theme(legend.position = "top",
            strip.text = element_text(face = "bold"))+ geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9)
    
    print(p)
  })
```


# Plot top 10 Metabolites and FC_0w, FC_3w and LogFC12w

```{r}
all_top_MElaSJR1000_df<-filter(all_top_MElaSJR1000_df, organ=="Heart_D")
for (organ in unique(all_top_MElaSJR1000_df$organ)) {
  temp <- all_top_MElaSJR1000_df[all_top_MElaSJR1000_df$organ == organ,]
  
  long_noflatData_with_time_top <- pivot_longer(temp, cols = c(FC_0w, FC_3w, FC_12w, FC_24w), 
                          names_to = "Time", values_to = "LogFC")
  long_noflatData_with_time_top$Time <- factor(long_noflatData_with_time_top$Time, levels=c("FC_0w", "FC_3w", "FC_12w", "FC_24w"))
  
  for (metabolite in unique(long_noflatData_with_time_top$ID)) {
    temp<-long_noflatData_with_time_top[long_noflatData_with_time_top$ID==metabolite, ]
    p=ggplot(temp, aes(x = Time, y = LogFC, group=1)) + geom_line() + geom_point() + labs(title = paste0("Top MElas: ", metabolite, " FC Between Endpoints in ", organ), x = "Endpoint (weeks)", y = "FC")+ theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(2.8))) + theme(axis.title.x = element_text(size = rel(1.8))) +theme(axis.title.x = element_blank(), axis.text.x = element_blank(), title= element_blank())
  print(p)
  plotname2 <- paste0("images2/", metabolite, "lineplot.png")
  ggsave(plotname2, dpi = 300) 

  }
}
```

# Plot Bottom 10 Metabolites and FC_0w, FC_3w and LogFC12w
```{r}
all_bottom_MElaSJR1000_df<-filter(all_bottom_MElaSJR1000_df, organ=="Heart_D")
for (organ in unique(all_bottom_MElaSJR1000_df$organ)) {
  temp <- all_bottom_MElaSJR1000_df[all_bottom_MElaSJR1000_df$organ == organ,]
  
  long_noflatData_with_time_bottom <- pivot_longer(temp, cols = c(FC_0w, FC_3w, FC_12w, FC_24w), 
                          names_to = "Time", values_to = "LogFC")
  long_noflatData_with_time_bottom$Time <- factor(long_noflatData_with_time_bottom$Time, levels=c("FC_0w", "FC_3w", "FC_12w", "FC_24w"))
  
  for (metabolite in unique(long_noflatData_with_time_bottom$ID)) {
    temp<-long_noflatData_with_time_bottom[long_noflatData_with_time_bottom$ID==metabolite, ]
    p=ggplot(temp, aes(x = Time, y = LogFC, group=1)) + geom_line() + geom_point() + labs(title = paste0("Bottom MElas: ", metabolite, " FC Between Endpoints in ", organ), x = "Endpoint Rations", y = "FC")+ theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(size=13)) + theme(axis.title.y = element_text(size = rel(2.8))) + theme(axis.title.x = element_text(size = rel(1.8))) +theme(axis.title.x = element_blank(), axis.text.x = element_blank(), title= element_blank())
  print(p)
  plotname2 <- paste0("images2/", metabolite, "lineplot_bottom.png")
  ggsave(plotname2, dpi = 300)
  }
}

```

#JR, 100 Analysis (make sure it looks the same as JR 1000 code!!!!)

```{r}
#Top MElaS
noflatDatalist = list()
for (organ in unique(JR_100_noflatData$organ)) {
  temp <- JR_100_noflatData[JR_100_noflatData$organ == organ,]
  
  top<-top_n(temp, 10, MElaS) #metabolites with top MElaS scores
  
  long_noflatData <- pivot_longer(top, cols = c(logFC_3v0, logFC_12v3),names_to = "Ratio", values_to = "LogFC")
  long_noflatData$Ratio <- gsub("Ratio", "", long_noflatData$Ratio)
  p=ggplot(long_noflatData, aes(x = Ratio, y = LogFC, group = ID, color=ID)) + geom_line(alpha = 0.6) + geom_point() + labs(title = paste0("Top MElas LogFC Between LogFC3v0 and LogFC12v3 in ", organ), "Top MElas LogFC Between LogFC3v0 and LogFC12v3 in ", x = "Post Infection Ratio", y = "LogFC")+ theme_minimal() 
  p+labs(color="Metabolite") +  theme(axis.text=element_text(size=10), axis.title=element_text(size=20),plot.title = element_text(size=22, hjust=0.5))+theme(legend.text = element_text(size = 10), legend.title = element_text(size = 15))
  print(p)
  noflatDatalist[[length(noflatDatalist) + 1]] <- top
}
all_top_MElaSJR100_df<-do.call(rbind, noflatDatalist)
```

```{r}
#Bottom MElaS
noflatDatalist2 = list()
for (organ in unique(JR_100_noflatData$organ)) {
  temp <- JR_100_noflatData[JR_100_noflatData$organ == organ,]
  
  bottom<-top_n(temp, -10, MElaS) #metabolites with lowest MElaS scores
  bottom <- bottom[sample(nrow(bottom), 10), ] #randomly sample 10
  long_noflatData <- pivot_longer(bottom, cols = c(logFC_3v0, logFC_12v3), 
                          names_to = "Ratio", values_to = "LogFC")
  long_noflatData$Ratio <- gsub("Ratio", "", long_noflatData$Ratio)
  p=ggplot(long_noflatData, aes(x = Ratio, y = LogFC, group = ID, color=ID)) +
  geom_line(alpha = 0.6) +
  geom_point() +
  labs(title = paste0("Bottom MElas LogFC Between LogFC3v0 and LogFC12v3 in ", organ),
       x = "Post Infection Ratio", y = "LogFC") +
  theme_minimal()
  p+labs(color="Metabolite") +  
		theme(axis.text=element_text(size=10),
				axis.title=element_text(size=20),plot.title = element_text(size=22, hjust=0.5))+theme(legend.text = element_text(size = 10), legend.title = element_text(size = 15))
  print(p)
  noflatDatalist2[[length(noflatDatalist2) + 1]] <- bottom

}
all_bottom_MElaSJR100_df<-do.call(rbind, noflatDatalist2)
```

```{r}
JR_100_metadata <- metadata %>%
  filter(ATTRIBUTE_parasite_species!="CL_Luc", ATTRIBUTE_infected_parasites!="1000") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='100' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

```

#boxplot comparing peak areas of each top MElaS metabolites by post infection days and condition
```{r}

plots <- all_top_MElaSJR100_df %>%
  split(.$organ) %>%
  imap(function(df, organ_name) {
    # Get the top metabolite names for this organ
    top_metabolites <- df$ID
    
    # Subset peak data for this organ and those metabolites
    subset_data <- JR_100_metadata  %>%
      filter(ATTRIBUTE_organ == organ_name) %>%
      select(ATTRIBUTE_infected_parasites, ATTRIBUTE_organ, ATTRIBUTE_endpoint, all_of(top_metabolites)) %>%
      pivot_longer(cols = all_of(top_metabolites),
                   names_to = "metabolite", values_to = "peak_area")
    
    # Make boxplots comparing timepoints
    p <- ggplot(subset_data, aes(x = ATTRIBUTE_endpoint, y = peak_area, fill = ATTRIBUTE_infected_parasites)) +
      geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
      facet_wrap(~ metabolite, scales = "free_y", ncol = 5) +
      theme_bw(base_size = 9) +
      labs(title = paste("Top 10 Metabolites in", organ_name, 'for JR, 100'),
           y = "Peak Area", x = "Timepoint", fill="Infection Status") +
      theme(legend.position = "top",
            strip.text = element_text(face = "bold"))+ geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9)
    
    print(p)
  })
```

#boxplot comparing peak areas of each bottom MElaS metabolites by post infection days and condition

```{r}
plots <- all_bottom_MElaSJR100_df %>%
  split(.$organ) %>%
  imap(function(df, organ_name) {
    # Get the top metabolite names for this organ
    bottom_metabolites <- df$ID
    
    # Subset peak data for this organ and those metabolites
    subset_data <- JR_100_metadata  %>%
      filter(ATTRIBUTE_organ == organ_name) %>%
      select(ATTRIBUTE_infected_parasites, ATTRIBUTE_organ, ATTRIBUTE_endpoint, all_of(bottom_metabolites)) %>%
      pivot_longer(cols = all_of(bottom_metabolites),
                   names_to = "metabolite", values_to = "peak_area")
    
    # Make boxplots comparing timepoints
    p <- ggplot(subset_data, aes(x = ATTRIBUTE_endpoint, y = peak_area, fill = ATTRIBUTE_infected_parasites)) +
      geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
      facet_wrap(~ metabolite, scales = "free_y", ncol = 5) +
      theme_bw(base_size = 9) +
      labs(title = paste("Bottom 10 Metabolites in", organ_name, 'for JR, 100'),
           y = "Peak Area", x = "Timepoint", fill="Infection Status") +
      theme(legend.position = "top",
            strip.text = element_text(face = "bold"))+ geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9)
    
    print(p)
  })
```

# Plot top 10 Metabolites and FC_0w, FC_3w and LogFC12w
```{r}
for (organ in unique(all_top_MElaSJR100_df$organ)) {
  temp <- all_top_MElaSJR100_df[all_top_MElaSJR100_df$organ == organ,]
  
  long_noflatData_with_time_top <- pivot_longer(temp, cols = c(FC_0w, FC_3w, FC_12w), 
                          names_to = "Time", values_to = "LogFC")
  long_noflatData_with_time_top$Time <- factor(long_noflatData_with_time_top$Time, levels=c("FC_0w", "FC_3w", "FC_12w"))
  
  p=ggplot(long_noflatData_with_time_top, aes(x = Time, y = LogFC, group = ID, color=ID)) + geom_line(alpha = 0.6) + geom_point() + labs(title = paste0("Top MElas LogFC Between LogFC3v0 and LogFC12v3 in ", organ), x = "Post Infection Ratio", y = "LogFC")+ theme_minimal() 
  p+labs(color="Metabolite") +  theme(axis.text=element_text(size=10), axis.title=element_text(size=20),plot.title = element_text(size=22, hjust=0.5))+theme(legend.text = element_text(size = 10), legend.title = element_text(size = 15))
  print(p)
}
```

# Plot Bottom 10 Metabolites and FC_0w, FC_3w and LogFC12w
```{r}

for (organ in unique(all_bottom_MElaSJR100_df$organ)) {
  temp <- all_bottom_MElaSJR100_df[all_bottom_MElaSJR100_df$organ == organ,]
  
  long_noflatData_with_time_bottom <- pivot_longer(temp, cols = c(FC_0w, FC_3w, FC_12w), 
                          names_to = "Time", values_to = "LogFC")
  long_noflatData_with_time_bottom$Time <- factor(long_noflatData_with_time_bottom$Time, levels=c("FC_0w", "FC_3w", "FC_12w"))
  
  p=ggplot(long_noflatData_with_time_bottom, aes(x = Time, y = LogFC, group = ID, color=ID)) + geom_line(alpha = 0.6) + geom_point() + labs(title = paste0("Bottom MElas LogFC Between LogFC3v0 and LogFC12v3 in ", organ), x = "Post Infection Ratio", y = "LogFC")+ theme_minimal() 
  p+labs(color="Metabolite") +  theme(axis.text=element_text(size=10), axis.title=element_text(size=20),plot.title = element_text(size=22, hjust=0.5))+theme(legend.text = element_text(size = 10), legend.title = element_text(size = 15))
  print(p)
}
```

# CL, 1000 Analysis

```{r}
#Top MElaS
noflatDatalist = list()
for (organ in unique(CL_1000_noflatData$organ)) {
  temp <- CL_1000_noflatData[CL_1000_noflatData$organ == organ,]
  
  top<-top_n(temp, 10, MElaS) #metabolites with top MElaS scores
  
  long_noflatData <- pivot_longer(top, cols = c(logFC_3v0, logFC_12v3),names_to = "Ratio", values_to = "LogFC")
  long_noflatData$Ratio <- gsub("Ratio", "", long_noflatData$Ratio)
  p=ggplot(long_noflatData, aes(x = Ratio, y = LogFC, group = ID, color=ID)) + geom_line(alpha = 0.6) + geom_point() + labs(title = paste0("Top MElas LogFC Between LogFC3v0 and LogFC12v3 in ", organ), "Top MElas LogFC Between LogFC3v0 and LogFC12v3 in ", x = "Post Infection Ratio", y = "LogFC")+ theme_minimal() 
  p+labs(color="Metabolite") +  theme(axis.text=element_text(size=10), axis.title=element_text(size=20),plot.title = element_text(size=22, hjust=0.5))+theme(legend.text = element_text(size = 10), legend.title = element_text(size = 15))
  print(p)
  noflatDatalist[[length(noflatDatalist) + 1]] <- top
}
all_top_MElaSCL1000_df<-do.call(rbind, noflatDatalist)
```

```{r}
#Bottom MElaS
noflatDatalist2 = list()
for (organ in unique(CL_1000_noflatData$organ)) {
  temp <- CL_1000_noflatData[CL_1000_noflatData$organ == organ,]
  
  bottom<-top_n(temp, -10, MElaS) #metabolites with lowest MElaS scores
  bottom <- bottom[sample(nrow(bottom), 10), ] #randomly sample 10
  long_noflatData <- pivot_longer(bottom, cols = c(logFC_3v0, logFC_12v3), 
                          names_to = "Ratio", values_to = "LogFC")
  long_noflatData$Ratio <- gsub("Ratio", "", long_noflatData$Ratio)
  p=ggplot(long_noflatData, aes(x = Ratio, y = LogFC, group = ID, color=ID)) +
  geom_line(alpha = 0.6) +
  geom_point() +
  labs(title = paste0("Bottom MElas LogFC Between LogFC3v0 and LogFC12v3 in ", organ),
       x = "Post Infection Ratio", y = "LogFC") +
  theme_minimal()
  p+labs(color="Metabolite") +  
		theme(axis.text=element_text(size=10),
				axis.title=element_text(size=20),plot.title = element_text(size=22, hjust=0.5))+theme(legend.text = element_text(size = 10), legend.title = element_text(size = 15))
  print(p)
  noflatDatalist2[[length(noflatDatalist2) + 1]] <- bottom

}
all_bottom_MElaSCL1000_df<-do.call(rbind, noflatDatalist2)
```


```{r}
CL_1000_metadata <- metadata %>%
  filter(ATTRIBUTE_parasite_species!="JR_Luc", ATTRIBUTE_infected_parasites!="100") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='1000' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

```

#boxplot comparing peak areas of each top MElaS metabolites by post infection days and condition
```{r}

plots <- all_top_MElaSCL1000_df %>%
  split(.$organ) %>%
  imap(function(df, organ_name) {
    # Get the top metabolite names for this organ
    top_metabolites <- df$ID
    
    # Subset peak data for this organ and those metabolites
    subset_data <- CL_1000_metadata  %>%
      filter(ATTRIBUTE_organ == organ_name) %>%
      select(ATTRIBUTE_infected_parasites, ATTRIBUTE_organ, ATTRIBUTE_endpoint, all_of(top_metabolites)) %>%
      pivot_longer(cols = all_of(top_metabolites),
                   names_to = "metabolite", values_to = "peak_area")
    
    # Make boxplots comparing timepoints
    p <- ggplot(subset_data, aes(x = ATTRIBUTE_endpoint, y = peak_area, fill = ATTRIBUTE_infected_parasites)) +
      geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
      facet_wrap(~ metabolite, scales = "free_y", ncol = 5) +
      theme_bw(base_size = 9) +
      labs(title = paste("Top 10 Metabolites in", organ_name, 'for CL, 1000'),
           y = "Peak Area", x = "Timepoint", fill="Infection Status") +
      theme(legend.position = "top",
            strip.text = element_text(face = "bold"))+ geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9)
    
    print(p)
  })
```

#boxplot comparing peak areas of each bottom MElaS metabolites by post infection days and condition

```{r}
plots <- all_bottom_MElaSCL1000_df %>%
  split(.$organ) %>%
  imap(function(df, organ_name) {
    # Get the top metabolite names for this organ
    bottom_metabolites <- df$ID
    
    # Subset peak data for this organ and those metabolites
    subset_data <- CL_1000_metadata  %>%
      filter(ATTRIBUTE_organ == organ_name) %>%
      select(ATTRIBUTE_infected_parasites, ATTRIBUTE_organ, ATTRIBUTE_endpoint, all_of(bottom_metabolites)) %>%
      pivot_longer(cols = all_of(bottom_metabolites),
                   names_to = "metabolite", values_to = "peak_area")
    
    # Make boxplots comparing timepoints
    p <- ggplot(subset_data, aes(x = ATTRIBUTE_endpoint, y = peak_area, fill = ATTRIBUTE_infected_parasites)) +
      geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
      facet_wrap(~ metabolite, scales = "free_y", ncol = 5) +
      theme_bw(base_size = 9) +
      labs(title = paste("Bottom 10 Metabolites in", organ_name, 'for CL, 1000'),
           y = "Peak Area", x = "Timepoint", fill="Infection Status") +
      theme(legend.position = "top",
            strip.text = element_text(face = "bold"))+ geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9)
    
    print(p)
  })
```

# Plot top 10 Metabolites and FC_0w, FC_3w and LogFC12w
```{r}
for (organ in unique(all_top_MElaSCL1000_df$organ)) {
  temp <- all_top_MElaSCL1000_df[all_top_MElaSCL1000_df$organ == organ,]
  
  long_noflatData_with_time_top <- pivot_longer(temp, cols = c(FC_0w, FC_3w, FC_12w), 
                          names_to = "Time", values_to = "LogFC")
  long_noflatData_with_time_top$Time <- factor(long_noflatData_with_time_top$Time, levels=c("FC_0w", "FC_3w", "FC_12w"))
  
  p=ggplot(long_noflatData_with_time_top, aes(x = Time, y = LogFC, group = ID, color=ID)) + geom_line(alpha = 0.6) + geom_point() + labs(title = paste0("Top MElas LogFC Between LogFC3v0 and LogFC12v3 in ", organ), x = "Post Infection Ratio", y = "LogFC")+ theme_minimal() 
  p+labs(color="Metabolite") +  theme(axis.text=element_text(size=10), axis.title=element_text(size=20),plot.title = element_text(size=22, hjust=0.5))+theme(legend.text = element_text(size = 10), legend.title = element_text(size = 15))
  print(p)
}
```

# Plot Bottom 10 Metabolites and FC_0w, FC_3w and LogFC12w
```{r}
for (organ in unique(all_bottom_MElaSCL1000_df$organ)) {
  temp <- all_bottom_MElaSCL1000_df[all_bottom_MElaSCL1000_df$organ == organ,]
  
  long_noflatData_with_time_bottom <- pivot_longer(temp, cols = c(FC_0w, FC_3w, FC_12w), 
                          names_to = "Time", values_to = "LogFC")
  long_noflatData_with_time_bottom$Time <- factor(long_noflatData_with_time_top$Time, levels=c("FC_0w", "FC_3w", "FC_12w"))
  
  p=ggplot(long_noflatData_with_time_bottom, aes(x = Time, y = LogFC, group = ID, color=ID)) + geom_line(alpha = 0.6) + geom_point() + labs(title = paste0("Bottom MElas LogFC Between LogFC3v0 and LogFC12v3 in ", organ), x = "Post Infection Ratio", y = "LogFC")+ theme_minimal() 
  p+labs(color="Metabolite") +  theme(axis.text=element_text(size=10), axis.title=element_text(size=20),plot.title = element_text(size=22, hjust=0.5))+theme(legend.text = element_text(size = 10), legend.title = element_text(size = 15))
  print(p)
}
```

# CL, 100 Analysis

```{r}
#Top MElaS
noflatDatalist = list()
for (organ in unique(CL_100_noflatData$organ)) {
  temp <- CL_100_noflatData[CL_100_noflatData$organ == organ,]
  
  top<-top_n(temp, 10, MElaS) #metabolites with top MElaS scores
  
  long_noflatData <- pivot_longer(top, cols = c(logFC_3v0, logFC_12v3),names_to = "Ratio", values_to = "LogFC")
  long_noflatData$Ratio <- gsub("Ratio", "", long_noflatData$Ratio)
  p=ggplot(long_noflatData, aes(x = Ratio, y = LogFC, group = ID, color=ID)) + geom_line(alpha = 0.6) + geom_point() + labs(title = paste0("Top MElas LogFC Between LogFC3v0 and LogFC12v3 in ", organ), "Top MElas LogFC Between LogFC3v0 and LogFC12v3 in ", x = "Post Infection Ratio", y = "LogFC")+ theme_minimal() 
  p+labs(color="Metabolite") +  theme(axis.text=element_text(size=10), axis.title=element_text(size=20),plot.title = element_text(size=22, hjust=0.5))+theme(legend.text = element_text(size = 10), legend.title = element_text(size = 15))
  print(p)
  noflatDatalist[[length(noflatDatalist) + 1]] <- top
}
all_top_MElaSCL100_df<-do.call(rbind, noflatDatalist)
```

```{r}
#Bottom MElaS
noflatDatalist2 = list()
for (organ in unique(CL_100_noflatData$organ)) {
  temp <- CL_100_noflatData[CL_100_noflatData$organ == organ,]
  
  bottom<-top_n(temp, -10, MElaS) #metabolites with lowest MElaS scores
  bottom <- bottom[sample(nrow(bottom), 10), ] #randomly sample 10
  long_noflatData <- pivot_longer(bottom, cols = c(logFC_3v0, logFC_12v3), 
                          names_to = "Ratio", values_to = "LogFC")
  long_noflatData$Ratio <- gsub("Ratio", "", long_noflatData$Ratio)
  p=ggplot(long_noflatData, aes(x = Ratio, y = LogFC, group = ID, color=ID)) +
  geom_line(alpha = 0.6) +
  geom_point() +
  labs(title = paste0("Bottom MElas LogFC Between LogFC3v0 and LogFC12v3 in ", organ),
       x = "Post Infection Ratio", y = "LogFC") +
  theme_minimal()
  p+labs(color="Metabolite") +  
		theme(axis.text=element_text(size=10),
				axis.title=element_text(size=20),plot.title = element_text(size=22, hjust=0.5))+theme(legend.text = element_text(size = 10), legend.title = element_text(size = 15))
  print(p)
  noflatDatalist2[[length(noflatDatalist2) + 1]] <- bottom

}
all_bottom_MElaSCL100_df<-do.call(rbind, noflatDatalist2)
```

```{r}
CL_100_metadata <- metadata %>%
  filter(ATTRIBUTE_parasite_species!="JR_Luc", ATTRIBUTE_infected_parasites!="1000") %>%
  mutate(ATTRIBUTE_infected_parasites=case_when(ATTRIBUTE_infected_parasites == '0'  ~ 'uninfected', ATTRIBUTE_infected_parasites=='100' ~ 'infected')) %>%
    mutate(ATTRIBUTE_endpoint = factor(ATTRIBUTE_endpoint, levels = c("0w", "3w", "12w")), ATTRIBUTE_infected_parasites=factor(ATTRIBUTE_infected_parasites, levels=c('uninfected', 'infected')))

```


#boxplot comparing peak areas of each top MElaS metabolites by post infection days and condition
```{r}

plots <- all_top_MElaSCL100_df %>%
  split(.$organ) %>%
  imap(function(df, organ_name) {
    # Get the top metabolite names for this organ
    top_metabolites <- df$ID
    
    # Subset peak data for this organ and those metabolites
    subset_data <- CL_100_metadata  %>%
      filter(ATTRIBUTE_organ == organ_name) %>%
      select(ATTRIBUTE_infected_parasites, ATTRIBUTE_organ, ATTRIBUTE_endpoint, all_of(top_metabolites)) %>%
      pivot_longer(cols = all_of(top_metabolites),
                   names_to = "metabolite", values_to = "peak_area")
    
    # Make boxplots comparing timepoints
    p <- ggplot(subset_data, aes(x = ATTRIBUTE_endpoint, y = peak_area, fill = ATTRIBUTE_infected_parasites)) +
      geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
      facet_wrap(~ metabolite, scales = "free_y", ncol = 5) +
      theme_bw(base_size = 9) +
      labs(title = paste("Top 10 Metabolites in", organ_name, 'for CL, 100'),
           y = "Peak Area", x = "Timepoint", fill="Infection Status") +
      theme(legend.position = "top",
            strip.text = element_text(face = "bold"))+ geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9)
    
    print(p)
  })
```

#boxplot comparing peak areas of each bottom MElaS metabolites by post infection days and condition

```{r}
plots <- all_bottom_MElaSCL100_df %>%
  split(.$organ) %>%
  imap(function(df, organ_name) {
    # Get the top metabolite names for this organ
    bottom_metabolites <- df$ID
    
    # Subset peak data for this organ and those metabolites
    subset_data <- CL_100_metadata  %>%
      filter(ATTRIBUTE_organ == organ_name) %>%
      select(ATTRIBUTE_infected_parasites, ATTRIBUTE_organ, ATTRIBUTE_endpoint, all_of(bottom_metabolites)) %>%
      pivot_longer(cols = all_of(bottom_metabolites),
                   names_to = "metabolite", values_to = "peak_area")
    
    # Make boxplots comparing timepoints
    p <- ggplot(subset_data, aes(x = ATTRIBUTE_endpoint, y = peak_area, fill = ATTRIBUTE_infected_parasites)) +
      geom_boxplot(position = position_dodge(width = 0.8), outlier.shape = NA) +
      facet_wrap(~ metabolite, scales = "free_y", ncol = 5) +
      theme_bw(base_size = 9) +
      labs(title = paste("Bottom 10 Metabolites in", organ_name, 'for CL, 100'),
           y = "Peak Area", x = "Timepoint", fill="Infection Status") +
      theme(legend.position = "top",
            strip.text = element_text(face = "bold"))+ geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                  alpha = 0.5, size = 0.9)
    
    print(p)
  })
```

# Plot top 10 Metabolites and FC_0w, FC_3w and LogFC12w
```{r}
for (organ in unique(all_top_MElaSCL100_df$organ)) {
  temp <- all_top_MElaSCL100_df[all_top_MElaSCL100_df$organ == organ,]
  
  long_noflatData_with_time_top <- pivot_longer(temp, cols = c(FC_0w, FC_3w, FC_12w), 
                          names_to = "Time", values_to = "LogFC")
  long_noflatData_with_time_top$Time <- factor(long_noflatData_with_time_top$Time, levels=c("FC_0w", "FC_3w", "FC_12w"))
  
  p=ggplot(long_noflatData_with_time_top, aes(x = Time, y = LogFC, group = ID, color=ID)) + geom_line(alpha = 0.6) + geom_point() + labs(title = paste0("Top MElas LogFC Between LogFC3v0 and LogFC12v3 in ", organ), x = "Post Infection Ratio", y = "LogFC")+ theme_minimal() 
  p+labs(color="Metabolite") +  theme(axis.text=element_text(size=10), axis.title=element_text(size=20),plot.title = element_text(size=22, hjust=0.5))+theme(legend.text = element_text(size = 10), legend.title = element_text(size = 15))
  print(p)
}
```

# Plot Bottom 10 Metabolites and FC_0w, FC_3w and LogFC12w
```{r}
for (organ in unique(all_bottom_MElaSCL100_df$organ)) {
  temp <- all_bottom_MElaSCL100_df[all_bottom_MElaSCL100_df$organ == organ,]
  
  long_noflatData_with_time_bottom <- pivot_longer(temp, cols = c(FC_0w, FC_3w, FC_12w), 
                          names_to = "Time", values_to = "LogFC")
  long_noflatData_with_time_bottom$Time <- factor(long_noflatData_with_time_top$Time, levels=c("FC_0w", "FC_3w", "FC_12w"))
  
  p=ggplot(long_noflatData_with_time_bottom, aes(x = Time, y = LogFC, group = ID, color=ID)) + geom_line(alpha = 0.6) + geom_point() + labs(title = paste0("Bottom MElas LogFC Between LogFC3v0 and LogFC12v3 in ", organ), x = "Post Infection Ratio", y = "LogFC")+ theme_minimal() 
  p+labs(color="Metabolite") +  theme(axis.text=element_text(size=10), axis.title=element_text(size=20),plot.title = element_text(size=22, hjust=0.5))+theme(legend.text = element_text(size = 10), legend.title = element_text(size = 15))
  print(p)
}
```

# Boxplots comparing parasite load and species for all organs (can use next chunk to make boxplots)
```{r}

for (organ in unique(noflatData$organ)) {

  temp <- noflatData[noflatData$organ == organ, ]

  # Create combined category for coloring
  temp$strain_conc <- paste0(temp$parasite_species, "_", temp$infected_parasites)

  # Define colors
  fill_colors <- c(
    "CL_100"  = "#55DDE0",  # light blue
    "CL_1000" = "#33658A",  # dark blue
    "JR_100"  = "#FFB4B4",  # light red
    "JR_1000" = "#950606"   # dark red
  )

  plot <- ggplot(temp, aes(
      x = parasite_species,
      y = MElaS,
      fill = strain_conc
    )) +
    geom_boxplot() +
    stat_compare_means(aes(group = parasite_species),
                       method = "wilcox.test",
                       label = "p.signif",
                       label.x = 1.5,
                       label.y = 17) +
    stat_compare_means(aes(group = infected_parasites),
                       method = "wilcox.test",
                       label = "p.signif",
                       label.x = 10,
                       label.y = 10) +
    scale_fill_manual(values = fill_colors,
                      name = "Strain and Load") +
    ggtitle(paste0("Boxplot Comparing MElaS by Parasite Inoculation \n    and Parasite Strain in the ", organ)) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
      axis.text.y = element_text(size = 13),
      axis.title.x = element_text(size = rel(1.8)),
      axis.title.y = element_text(size = rel(1.8)),
      legend.text = element_text(size = 13),
      legend.title = element_text(size = rel(1.8)),
      plot.title = element_text(size = 20)
    ) +
    labs(x = "Parasite Strain", y = "MElaS")

  print(plot)
}
```


# p-values comparing CL vs JR at 100 and 1000 and boxplots Works!!!
```{r}
noflatData$Group <- paste0(noflatData$parasite_species, "_", noflatData$infected_parasites)

p_values_df <- data.frame(Organ = character(),
                          Group1 = character(),
                          Group2 = character(),
                          Test = character(),
                          P_Value = numeric(),
                          Fold_Change = numeric(),
                          stringsAsFactors = FALSE)

OrganList <- c(unique(noflatData$organ))
for (Organ in OrganList){
  print(Organ)
  temp <- noflatData[noflatData$organ == Organ,]
  comparisons <- combn(levels(factor(temp$Group)), 2, simplify = FALSE)
  comparisons <- c(comparisons[2], comparisons[5])
    for (comp in comparisons) {
      group1 <- comp[1]
      group2 <- comp[2]
      
      # Subset data for the two groups
      temp_subset <- temp[temp$Group %in% c(group1, group2), ]
      
      if (length(unique(temp_subset[temp_subset$Group == group1,]$MElaS)) > 1 & length(temp_subset[temp_subset$Group == group1,]$MElaS) > 2) {
        normalcy1 <- shapiro.test(temp_subset[temp_subset$Group == group1,]$MElaS)$p.value
        } else {
          normalcy1 <- 0 # Force Wilcoxon test if no variability
        }
      if (length(unique(temp_subset[temp_subset$Group == group2,]$MElaS)) > 1 & length(temp_subset[temp_subset$Group == group2,]$MElaS) > 2) {
        normalcy2 <- shapiro.test(temp_subset[temp_subset$Group == group2,]$MElaS)$p.value
        } else {
          normalcy2 <- 0 # Force Wilcoxon test if no variability
        }
    
      # Choose the appropriate test based on normality
      if (normalcy1 < 0.05 | normalcy2 < 0.05) {
        test_used <- "Wilcoxon"
        plotTest <- "wilcox.test"
        p_val <- wilcox.test(MElaS ~ Group, data = temp_subset)$p.value
      } else {
        test_used <- "T-test"
        plotTest <- "t.test"
        p_val <- t.test(MElaS ~ Group, data = temp_subset)$p.value
      }
      
      
      fc <- median(temp_subset[temp_subset$Group == group1,]$MElaS)/median(temp_subset[temp_subset$Group == group2,]$MElaS)
      
      # Store results in the data frame
      p_values_df <- rbind(p_values_df, data.frame(
        Organ = Organ,
        Group1 = group1,
        Group2 = group2,
        Test = test_used,
        Fold_Change = fc,
        P_Value = p_val
      ))
    }
  fill_colors <- c(
    "CL_100"  = "#55DDE0",  # light blue
    "CL_1000" = "#33658A",  # dark blue
    "JR_100"  = "#FFB4B4",  # light red
    "JR_1000" = "#950606"   # dark red
  )
  plot <- ggplot(temp, aes(x = parasite_species, y = MElaS, fill = parasite_species)) +
    facet_wrap(~ infected_parasites) +
    geom_boxplot(outlier.shape = NA) +
      stat_compare_means(aes(group = parasite_species), 
                         method = plotTest, 
                         label = "p.signif",
                         label.x = 1.35) +
    labs(
      title = paste0(Organ),
      x = "Parasite Strain",
      y = "MElaS",
      fill = "Parasite Strain"
    ) +
    scale_fill_manual(values = c("#55DDE0", "#FFB4B4")) +
    theme_bw() +
        theme(axis.title.x = element_blank(),           plot.title = element_text(hjust = 0.5))
    print(plot)
    #plotname <- paste0("Esophagus_", molecule, ".png")
    #ggsave(plotname, dpi = 300)
    
    
}

#write.csv(p_values_df, "by_parasite_group_pvalues.csv", row.names = FALSE)
```

# p-values comparing 100 vs 1000 at JR and CL and boxplots Works!!!

```{r}
p_values_df_2 <- data.frame(Organ = character(),
                          Group1 = character(),
                          Group2 = character(),
                          Test = character(),
                          Fold_Change = numeric(),
                          P_Value = numeric(),
                          stringsAsFactors = FALSE)

OrganList <- c(unique(noflatData$organ))
for (Organ in OrganList){
  print(Organ)
  temp <- noflatData[noflatData$organ == Organ,]
  comparisons <- combn(levels(factor(temp$Group)), 2, simplify = FALSE)
  comparisons <- c(comparisons[1], comparisons[6])
    for (comp in comparisons) {
      group1 <- comp[1]
      group2 <- comp[2]
      
      # Subset data for the two groups
      temp_subset <- temp[temp$Group %in% c(group1, group2), ]
      
      if (length(unique(temp_subset[temp_subset$Group == group1,]$MElaS)) > 1 & length(temp_subset[temp_subset$Group == group1,]$MElaS) > 2) {
        normalcy1 <- shapiro.test(temp_subset[temp_subset$Group == group1,]$MElaS)$p.value
        } else {
          normalcy1 <- 0 # Force Wilcoxon test if no variability
        }
      if (length(unique(temp_subset[temp_subset$Group == group2,]$MElaS)) > 1 & length(temp_subset[temp_subset$Group == group2,]$MElaS) > 2) {
        normalcy2 <- shapiro.test(temp_subset[temp_subset$Group == group2,]$MElaS)$p.value
        } else {
          normalcy2 <- 0 # Force Wilcoxon test if no variability
        }
    
      # Choose the appropriate test based on normality
      if (normalcy1 < 0.05 | normalcy2 < 0.05) {
        test_used <- "Wilcoxon"
        plotTest <- "wilcox.test"
        p_val <- wilcox.test(MElaS ~ Group, data = temp_subset)$p.value
      } else {
        test_used <- "T-test"
        plotTest <- "t.test"
        p_val <- t.test(MElaS ~ Group, data = temp_subset)$p.value
      }
      
      
      fc <- median(temp_subset[temp_subset$Group == group1,]$MElaS)/median(temp_subset[temp_subset$Group == group2,]$MElaS)
      
      # Store results in the data frame
      p_values_df_2 <- rbind(p_values_df_2, data.frame(
        Organ = Organ,
        Group1 = group1,
        Group2 = group2,
        Test = test_used,
        Fold_Change = fc,
        P_Value = p_val
      ))
    }
    
}
write.csv(p_values_df_2, "by_parasite_conc_pvalues.csv", row.names = FALSE)
```
#Proportion of non zeros vs zero MElaS scores

```{r}
#total percentage of nonzeros
non_zero_count=sum(data$MElaS !=0)
total_rows=nrow(data)
proportion_non_zero <- (non_zero_count / total_rows) *100
zero_count=total_rows-non_zero_count

#percentage of nonzeros by organ 
proportion_non_zero_by_organ <- data %>%
  group_by(organ, parasite_species, infected_parasites) %>%
  summarise(
    non_zero_count_organ=sum(MElaS !=0),
    total_rows_organ=n(),
    proportion_non_zero_organ = (non_zero_count_organ / total_rows_organ) *100,
    proportion_non_zero_organ =format(round(proportion_non_zero_organ, 3), nsmall = 3),
    zero_count=total_rows_organ-non_zero_count_organ
  )


proportion_non_zero_by_organ$proportion_non_zero_organ=as.numeric(proportion_non_zero_by_organ$proportion_non_zero_organ)
```

#Percentage of Non-Zero MElas Scores per Organ Barplot including JR, CL, 100 and 1000

```{r}
proportion_non_zero_by_organ$strain_conc <- paste0(
  proportion_non_zero_by_organ$parasite_species, "_",
  proportion_non_zero_by_organ$infected_parasites
)

# Colors
fill_colors <- c(
  "CL_100"  = "lightblue", 
  "CL_1000" = "blue3",
  "JR_100"  = "firebrick1",
  "JR_1000" = "red3"
)

ggplot(data = proportion_non_zero_by_organ,
       aes(x = as.factor(organ),
           y = proportion_non_zero_organ,
           fill = strain_conc)) +
  geom_col(position = position_dodge(width = 0.9)) + 
  facet_wrap(vars(infected_parasites)) +
  
  scale_fill_manual(values = fill_colors,
                    name = "Strain and Load") +
  
  ggtitle("Percentage of Non-Zero MElaS Scores") +
  
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.8)),
    axis.title.y = element_text(size = rel(1.8)),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = rel(1.8)),
    plot.title = element_text(size = 20)
  ) +
  labs(x = "Organ", y = "Percentage (%)")

```

#bar graphs comparing Percentage of Non-Zero MElas Scores by paraiste and load of each organ (bad)

```{r}
for (organ in unique(proportion_non_zero_by_organ$organ)) {
  temp <- proportion_non_zero_by_organ[proportion_non_zero_by_organ$organ == organ,]
  plot= ggplot(data=temp, aes(x=infected_parasites, y=proportion_non_zero_organ, fill=parasite_species)) +
  geom_bar(position='dodge', stat="identity")+
  geom_text(aes(label=proportion_non_zero_organ), vjust=3.3, size=2.5)+
  theme_minimal() +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle(paste0(organ, ": Percentage of Non-Zero MElas Scores")) + labs(x="infected_parasites", y="non-zero metabolite/total metabolite x 100")
  print(plot)
}

```

#Fisher's Test: comparing JR and CL at 100 and 1000 of each organ 
Fold change CL/JR at 1000 and CL/JR at 100

```{r}
fisher_test_results_byJRvCL <- data.frame(
  Organ = character(),
  P_Value = numeric(),
  Load = character(),
  Parasite = character(),
  FoldChange = numeric(),
  stringsAsFactors = FALSE
)

for (Tissue in unique(proportion_non_zero_by_organ$organ)) {
  proportion_non_zero_by_Organ=proportion_non_zero_by_organ %>%
  group_by(parasite_species, infected_parasites, organ) %>%
  filter(organ==Tissue)
  
  for (load in unique(proportion_non_zero_by_Organ$infected_parasites)) {
    noflatDataframe_infected_parasites=proportion_non_zero_by_Organ %>%
      filter(infected_parasites==load) %>%
      select(non_zero_count_organ, zero_count, proportion_non_zero_organ)
    print(noflatDataframe_infected_parasites)
    foldchange = noflatDataframe_infected_parasites$proportion_non_zero_organ[1] / noflatDataframe_infected_parasites$proportion_non_zero_organ[2]
    
    table_counts <- noflatDataframe_infected_parasites %>%
      group_by(parasite_species, infected_parasites) %>%
      summarise(non_zero = sum(non_zero_count_organ),
                zero = sum(zero_count)) %>%
      ungroup()
    
    contingency <- as.matrix(table_counts[, c("non_zero", "zero")])
    rownames(contingency) <- table_counts$parasite_species
    print(contingency)
    
    fisher_result <- fisher.test(contingency, workspace=2000000000)
    print(paste0("Fisher Test Comparing JR and CL at ", " Parasite Load in the ", organ))
    print(fisher_result)
    
    fisher_test_results_byJRvCL <- rbind(
      fisher_test_results_byJRvCL,
      data.frame(Load = load, p_value = fisher_result$p.value, Organ = Tissue, parasite=paste(rownames(contingency), collapse="_vs_"), FoldChange=foldchange)
    )

  }
}

fisher_test_results_byJRvCL_sig<-fisher_test_results_byJRvCL %>%
  filter(p_value<0.05)

write.csv(fisher_test_results_byJRvCL, "fisher_test_results_byJRvCL.csv", row.names = FALSE)
```

```{r}
ggplot(data=fisher_test_results_byJRvCL, aes(x=Organ, y=FoldChange, fill=Load)) +
  geom_bar(position='dodge', stat="identity")+
  theme_minimal() +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle("Fold Change per Organ") + labs(x="Organ", y="Fold Change")
```

#Fisher's Test: comparing 100 and 1000 at JR and CL of each organ
```{r}
fisher_test_results_by100v1000 <- data.frame(
  Organ = character(),
  P_Value = numeric(),
  Load = character(),
  parasite = character(),
  FoldChange = numeric(),
  stringsAsFactors = FALSE
)

for (Tissue in unique(proportion_non_zero_by_organ$organ)) {
  proportion_non_zero_by_Organ_100v1000=proportion_non_zero_by_organ %>%
  filter(organ==Tissue) %>%  
  group_by(parasite_species, infected_parasites, organ)
  
  for (Parasite in unique(proportion_non_zero_by_organ$parasite_species)) {
    dataframe_Parasite=proportion_non_zero_by_Organ_100v1000 %>%
      filter(parasite_species==Parasite) %>%
      select(non_zero_count_organ, zero_count, infected_parasites, proportion_non_zero_organ)
    #print(dataframe_Parasite)
    foldchange = dataframe_Parasite$proportion_non_zero_organ[1] / dataframe_Parasite$proportion_non_zero_organ[2]
    
    table_counts <- dataframe_Parasite %>%
      group_by(infected_parasites) %>%
      summarise(non_zero = sum(non_zero_count_organ),
                zero = sum(zero_count)) %>%
      ungroup()
    
    contingency <- as.matrix(table_counts[, c("non_zero", "zero")])
    rownames(contingency) <- table_counts$infected_parasites
    print(contingency)
    
    fisher_result <- fisher.test(contingency, workspace=2000000000)
    print(paste0("Fisher Test Comparing 100 and 1000 Parasite Load with in the ", Tissue))
    print(fisher_result)
    
    fisher_test_results_by100v1000 <- rbind(
      fisher_test_results_by100v1000,
      data.frame(Load = paste(rownames(contingency), collapse="_vs_"), P_value = fisher_result$p.value, Organ = Tissue, parasite_species=Parasite, FoldChange=foldchange)
    )

  }
}

#fisher_test_results_by100v1000_sig<-fisher_test_results_by100v1000 %>%
  #filter(P_value<0.05)

write.csv(fisher_test_results_by100v1000, "fisher_test_results_by100v1000.csv", row.names = FALSE)
```

# median values

```{r}
median_by_organ_infected_parasites_parasite <- aggregate(MElaS ~ organ + parasite_species+infected_parasites, data = noflatData, FUN = median, na.action = na.omit)

#write.csv(median_by_organ_infected_parasites_parasite,"median_by_organ_infected_parasites_parasite.csv")
median_by_organ_infected_parasites_parasite$organ <- factor(median_by_organ_infected_parasites_parasite$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(data=median_by_organ_infected_parasites_parasite, aes(x = as.factor(organ), y = MElaS, fill = as.factor(parasite_species))) +
  geom_col(position = "dodge") +
  facet_wrap(vars(infected_parasites)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle("Median MElas Scores per Organ") + labs(x="Organ", y="Median MElaS", fill="Parasite Species")

median_by_organ_infected_parasites_parasite$strain_conc <- paste0(
  median_by_organ_infected_parasites_parasite$parasite_species, "_",
  median_by_organ_infected_parasites_parasite$infected_parasites
)

# Colors
fill_colors <- c(
  "CL_100"  = "lightblue", 
  "CL_1000" = "blue3",
  "JR_100"  = "firebrick1",
  "JR_1000" = "red3"
)

noflatData$strain_conc <- paste0(
  noflatData$parasite_species, "_",
  noflatData$infected_parasites
)
ggplot(data=median_by_organ_infected_parasites_parasite, aes(x = as.factor(organ), y = MElaS, fill = strain_conc)) +
  geom_col(position = "dodge") +
  facet_wrap(vars(infected_parasites)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle("Median MElas Scores") + labs(x="Organ", y="Median MElaS") +
  
  scale_fill_manual(values = fill_colors,
                    name = "Strain and \nConcentration") +
  
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.8)),
    axis.title.y = element_text(size = rel(1.8)),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = rel(1.8)),
    plot.title = element_text(size = 20)
  )

median_by_organ_infected_parasites_parasite_1000 <- median_by_organ_infected_parasites_parasite %>%
  filter(infected_parasites == "1000")

ggplot(data=median_by_organ_infected_parasites_parasite_1000, aes(x = as.factor(organ), y = MElaS, fill = parasite_species)) +
  geom_col(position = "dodge") + geom_signif(comparisons = list(c("Cecum", "Distal_Colon")),
              map_signif_level = TRUE) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle("Median MElas Scores") + labs(x="Organ", y="Median MElaS") +
  
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.8)),
    axis.title.y = element_text(size = rel(1.8)),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = rel(1.8)),
    plot.title = element_text(size = 20)
  )

```

#boxplots comparing MElaS in all organs 

```{r}
#JR 1000

# do Kruskal Wallis test first (compare more than 3)
kruskalwalisJR1000 <- kruskal.test(MElaS ~ organ, data = JR_1000_noflatData)
kruskal.test(MElaS ~ organ, data = JR_1000_noflatData) #organs are significantly different p<2.2e-16

dunntestJR1000 <- dunnTest(MElaS ~ organ,
         data=JR_1000_noflatData,
         method="bonferroni")
dunn_df_JR1000 <- dunntestJR1000$res
dunn_df_JR1000_sig<-dunn_df_JR1000 %>%
  filter(P.adj<0.05)

JR_1000_noflatData$organ <- factor(JR_1000_noflatData$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(JR_1000_noflatData, aes(x=organ, y=MElaS, fill=organ))+ geom_boxplot() +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("MElaS of All Organs JR 1000") +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(plot.title=element_text(size=25)) +  labs(x="Organ", y="MElaS") + scale_fill_viridis(discrete = TRUE, option="C") +
    theme(legend.position="none")

#JR 100
kruskal.test(MElaS ~ organ, data = JR_100_noflatData) 
dunntestJR100 <- dunnTest(MElaS ~ organ,
         data=JR_100_noflatData,
         method="bonferroni")
dunn_df_JR100 <- dunntestJR100$res

JR_100_noflatData$organ <- factor(JR_100_noflatData$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(JR_100_noflatData, aes(x=organ, y=MElaS, fill=organ))+ geom_boxplot() +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("MElaS of All Organs JR 100") +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(plot.title=element_text(size=25)) +  labs(x="Organ", y="MElaS") + scale_fill_viridis(discrete = TRUE, option="C") +
    theme(legend.position="none")

#CL 1000
kruskal.test(MElaS ~ organ, data = CL_1000_noflatData) 
dunntestCL1000 <- dunnTest(MElaS ~ organ,
         data=CL_1000_noflatData,
         method="bonferroni")
dunn_df_CL1000 <- dunntestCL1000$res

CL_1000_noflatData$organ <- factor(CL_1000_noflatData$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(CL_1000_noflatData, aes(x=organ, y=MElaS, fill=organ))+ geom_boxplot() +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("MElaS of All Organs CL 1000") +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(plot.title=element_text(size=25)) +  labs(x="Organ", y="MElaS") + scale_fill_viridis(discrete = TRUE, option="C") +
    theme(legend.position="none")


#CL 100
kruskal.test(MElaS ~ organ, data = CL_100_noflatData) 
dunntestCL100 <-dunnTest(MElaS ~ organ,
         data=CL_100_noflatData,
         method="bonferroni", list = TRUE)
dunn_df_CL100 <- dunntestCL100$res

CL_100_noflatData$organ <- factor(CL_100_noflatData$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(CL_100_noflatData, aes(x=organ, y=MElaS, fill=organ))+ geom_boxplot() +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("MElaS of All Organs CL 100") +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(plot.title=element_text(size=25)) +  labs(x="Organ", y="MElaS") + scale_fill_viridis(discrete = TRUE, option="C") +
    theme(legend.position="none")


#all
noflatData$organ <- factor(noflatData$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(noflatData, aes(x=organ, y=MElaS, fill=organ))+ geom_boxplot() +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +ggtitle("MElaS of All Organs") +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(plot.title=element_text(size=25)) +  labs(x="Organ", y="MElaS") + scale_fill_viridis(discrete = TRUE, option="C") +
    theme(legend.position="none")

median_by_organ <- aggregate(MElaS ~ organ, data = noflatData, FUN = median, na.action = na.omit)

ggplot(data=median_by_organ, aes(x = as.factor(organ), y = MElaS)) +
  geom_col(position = "dodge") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle("Median MElas Scores per Organ") + labs(x="Organ", y="Median MElaS")

#write.csv(dunn_df_JR1000, "dunn_df_JR1000.csv", row.names = FALSE)
#write.csv(dunn_df_JR100, "dunn_df_JR100.csv", row.names = FALSE)
#write.csv(dunn_df_CL1000, "dunn_df_CL1000.csv", row.names = FALSE)
#write.csv(dunn_df_CL100, "dunn_df_CL100.csv", row.names = FALSE)

```

#comparing types
```{r}
dataWithMElaSTypesCount<-dataWithMElaSTypes %>%
  group_by(organ, parasite_species, infected_parasites, Type) %>%
  summarise(TypeCount = n())

dataWithMElaSTypes$Type <- factor(dataWithMElaSTypes$Type, levels=c("up_down", "down_up", "flat_up", "flat_down", "up_flat", "down_flat", "up_up", "down_down", "flat"))

noflatData$Type <- factor(noflatData$Type, levels=c("up_down", "down_up", "flat_up", "flat_down", "up_flat", "down_flat", "up_up", "down_down"))

dataWithMElaSTypesCountnoflat<-filter(dataWithMElaSTypesCount, Type!="flat", )

#counts of Types
ggplot(data=dataWithMElaSTypes, aes(x = Type)) + geom_bar() + ggtitle("Count of Types Including Flat") + labs(x="Type", y="Type Count")

dataWithMElaSTypesCountnoflat$Type <- factor(dataWithMElaSTypesCountnoflat$Type, levels=c("up_down", "down_up", "flat_up", "flat_down", "up_flat", "down_flat", "up_up", "down_down"))

ggplot(noflatData, aes(x = Type)) + geom_bar() + ggtitle("Count of Types") + labs(x="Type", y="Type Count")

#counts of type by parasite and concentration 
for (organ in unique(dataWithMElaSTypesCountnoflat$organ)) {
  temp <- dataWithMElaSTypesCountnoflat[dataWithMElaSTypesCountnoflat$organ == organ,]
  plot<-ggplot(data=temp, aes(x = as.factor(Type), y = TypeCount, fill = as.factor(parasite_species))) +
  geom_col(position = "dodge") +
  facet_wrap(vars(infected_parasites)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle(paste0("Count of Categories in ", organ)) + ylim(0, 700) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.8)),
    axis.title.y = element_text(size = rel(1.8)),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = rel(1.8)),
    plot.title = element_text(size = 20)
  ) +
  labs(x = "Organ", y = "Count", fill="Parasite Strain")
  print(plot)
}

dataWithMElaSTypesCountnoflat$strain_conc <- paste0(
  dataWithMElaSTypesCountnoflat$parasite_species, "_",
  dataWithMElaSTypesCountnoflat$infected_parasites
)


# Colors
fill_colors <- c(
  "CL_100"  = "lightblue", 
  "CL_1000" = "blue3",
  "JR_100"  = "firebrick1",
  "JR_1000" = "red3"
)

for (organ in unique(dataWithMElaSTypesCountnoflat$organ)) {
  temp <- dataWithMElaSTypesCountnoflat[dataWithMElaSTypesCountnoflat$organ == organ,]
  plot<-ggplot(data=temp, aes(x = as.factor(Type), y = TypeCount, fill = strain_conc)) +
  geom_col(position = "dodge") +
  facet_wrap(vars(infected_parasites)) + scale_fill_manual(values = fill_colors,
                    name = "Strain and Load")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ggtitle(paste0("Count of Categories in ", organ)) + ylim(0, 700) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.8)),
    axis.title.y = element_text(size = rel(1.8)),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = rel(1.8)),
    plot.title = element_text(size = 20)
  ) +
  labs(x = "Organ", y = "Count")
  print(plot)
}
#max MElaS by type barplot
ggplot(data=noflatData, aes(x = as.factor(Type), y=MElaS))+geom_col(position = "dodge") + ggtitle("Max MElas per Type w/out flat") + labs(x="Type", y="MElaS")

#max MElaS by type barplot with flat
ggplot(data=dataWithMElaSTypes, aes(x = as.factor(Type), y=MElaS))+geom_col(position = "dodge") + ggtitle("Max MElas per Type w/ flat") + labs(x="Type", y="MElaS") #higher values of flat

median_MElaS_per_Type <- noflatData %>%
  group_by(Type) %>%
  summarise(median_value = median(MElaS))

median_MElaS_per_Type$Type <- factor(median_MElaS_per_Type$Type, levels=c("up_down", "down_up", "flat_up", "flat_down", "up_flat", "down_flat", "up_up", "down_down"))


#median MElaS by type barplot
ggplot(median_MElaS_per_Type, aes(x = Type, y = median_value)) +
  geom_bar(stat = "identity") +
  labs(title = "Median MELaS by Category", x = "Category", y = "Median MElaS") +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(plot.title=element_text(size=30))


#IMPORTANT
ggplot(dataWithMElaSTypes, aes(x=Type, y=MElaS))+ geom_boxplot() + labs(title = "Boxplot Comparing MelaS of Each Category", x = "Category", y = "MElaS") +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=13)) + theme(axis.title.y = element_text(size = rel(1.8))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(plot.title=element_text(size=25))

dataWithMElaSTypes %>%
   group_by(Type) %>% 
   summarise(
      mean = mean(MElaS, na.rm=TRUE), 
      sd=sd(MElaS), na.rm=TRUE,
      max_score = max(MElaS, na.rm = TRUE),
      min_score = min(MElaS, na.rm = TRUE))
```



# Molecular networking
```{r}
molecules<-read.csv("DiseaseToleranceGNPSResults.csv")

#molecules_df <- molecules[- grep("N/A", molecules$superclass),]
  
# to compare all molecules types of organs
merged_all_df <- merge(
  molecules,
  dataWithMElaSTypes,
  by.x = "X.Scan.",
  by.y = "extracted_ID",
  all = FALSE   # inner join (only matching rows)
)

merged_all_df <- merged_all_df[, c( 'X.Scan.', 'superclass', 'class', 'ID', 'MElaS', 'logFC_3v0', 'logFC_12v3', 'FC_0w', 'FC_3w', 'FC_12w', 'FC_24w', 'organ', 'parasite_species', 'infected_parasites', 'Type', 'Type_24w')]

for (Organ in unique(merged_all_df$organ)) {
  temp<- merged_all_df %>%
    filter(organ %in% c(Organ))
  temp <- temp[- grep("N/A", temp$superclass),]
  plot <- ggplot(temp, aes(x = superclass)) +
  geom_bar() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5)) + labs(title = paste0("All ", Organ))
  print(plot)
  new_name <- paste0("df_", Organ)
  assign(new_name, temp)
}

# to compare top metabolic molecules types of organs
merged_all_top_df <- merge(
  molecules,
  all_top_df,
  by.x = "X.Scan.",
  by.y = "extracted_ID",
  all = FALSE   # inner join (only matching rows)
)

merged_all_top_df <- merged_all_top_df[, c( 'X.Scan.', 'superclass', 'class', 'ID', 'MElaS', 'logFC_3v0', 'logFC_12v3', 'FC_0w', 'FC_3w', 'FC_12w', 'FC_24w', 'organ', 'parasite_species', 'infected_parasites', 'Type', 'Type_24w')]

for (Organ in unique(merged_all_top_df$organ)) {
  top_temp<- merged_all_top_df %>%
    filter(organ %in% c(Organ))
  top_temp <- top_temp[- grep("N/A", top_temp$superclass),]
  plot1 <- ggplot(top_temp, aes(x = superclass)) +
  geom_bar() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5)) + labs(title = paste0("Top ", Organ))
  print(plot1)
}

merged_all_top_df_count<-merged_all_top_df %>%
  group_by(organ, superclass) %>%
  summarise(superclassCount = n())

# to compare bottom metabolic molecules types of organs
merged_all_bottom_df <- merge(
  molecules,
  all_bottom_df,
  by.x = "X.Scan.",
  by.y = "extracted_ID",
  all = FALSE   # inner join (only matching rows)
)

merged_all_bottom_df <- merged_all_bottom_df[, c( 'X.Scan.', 'superclass', 'class', 'ID', 'MElaS', 'logFC_3v0', 'logFC_12v3', 'FC_0w', 'FC_3w', 'FC_12w', 'FC_24w', 'organ', 'parasite_species', 'infected_parasites', 'Type', 'Type_24w')]

for (Organ in unique(merged_all_bottom_df$organ)) {
  bottom_temp<- merged_all_bottom_df %>%
    filter(organ %in% c(Organ))
  bottom_temp <- bottom_temp[- grep("N/A", bottom_temp$superclass),]
  plot2 <- ggplot(bottom_temp, aes(x = superclass)) +
  geom_bar() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5)) + labs(title = paste0("Bottom ", Organ))
  print(plot2)
}
merged_all_bottom_df_count<-merged_all_bottom_df %>%
  group_by(organ, superclass) %>%
  summarise(superclassCount = n())
write.csv(merged_all_bottom_df_count, "merged_all_bottom_df_count.csv")
write.csv(merged_all_top_df_count, "merged_all_top_df_count.csv")
```

# i dont know what this is...
```{r}

for (Organ in unique(all_bottom_MElaSJR1000_df$organ)) {
  bottom_temp <- all_bottom_MElaSJR1000_df %>%
    filter(organ %in% c(Organ)) %>%
    mutate(extracted_ID = str_remove_all(ID, "X(\\d+).(\\d+)_(\\d+).(\\d+)_"))
  all_bottom_MElaSJR1000_withmole <- filter(molecules_df, X.Scan. %in% bottom_temp$extracted_ID)
  all_bottom_MElaSJR1000_withmole <- all_bottom_MElaSJR1000_withmole[, c("X.Scan.", "Compound_Name", "MassDiff", "superclass", "class")]
  plot2 <- ggplot(all_bottom_MElaSJR1000_withmole, aes(x = superclass)) +
  geom_bar() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5)) + labs(title = paste0(organ, ": for bottom JR 1000" ))
  print(plot2)
}


for (Organ1 in unique(merged_df$organ)) {
    temp <- merged_df[merged_df$organ == Organ1,] 
    plot1 <- ggplot(temp, aes(x = superclass)) +
    geom_bar() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5)) + labs(title = paste0(Organ1, ": for all" ))
  print(plot1)
}

merged_df_cecum <- merged_df %>%
  filter(organ=="Cecum" )

ggplot(merged_df_cecum, aes(x = superclass)) +
    geom_bar() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))
  
merged_df_upperstom <- merged_df %>%
  filter(organ=="Upper_Stomach" )

ggplot(merged_df_upperstom, aes(x = superclass)) +
    geom_bar() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))
  
ggplot(all_top_MElaSJR1000_withmole, aes(x = superclass, y = count)) +
  geom_bar(stat = "identity") +
  labs(title = "Counts of type of molecule", x = "class", y = "count") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))

all_bottom_MElaSJR1000_withmole<-filter(all_bottom_MElaSJR1000_withmole, npclassifier_class!="N/A", )

all_bottom_MElaSJR1000_withmole<-all_bottom_MElaSJR1000_withmole %>%
  group_by(npclassifier_class) %>%
  mutate(count=n())
  
ggplot(all_bottom_MElaSJR1000_withmole, aes(x = npclassifier_class, y = count)) +
  geom_bar(stat = "identity") +
  labs(title = "Counts of type of molecule", x = "class", y = "count") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))

```

```{r}
noflatData <- noflatData %>%
  mutate(extracted_ID = str_remove_all(ID, "X(\\d+).(\\d+)_(\\d+).(\\d+)_"))

noflatData_withmole <- filter(molecules, X.Scan. %in% noflatData$extracted_ID)
noflatData_withmole <- noflatData_withmole[, c("X.Scan.", "Compound_Name", "MassDiff", "npclassifier_superclass", "npclassifier_class", "npclassifier_pathway")]

noflatData_withmole<-filter(noflatData_withmole, npclassifier_class!="N/A", )
noflatData_withmole<-filter(noflatData_withmole, npclassifier_superclass!="N/A", )

noflatData_withmole<-noflatData_withmole %>%
  group_by(npclassifier_class) %>%
  mutate(class=n())

noflatData_withmole<-noflatData_withmole %>%
  group_by(npclassifier_superclass) %>%
  mutate(superclass=n())

ggplot(noflatData_withmole, aes(x = npclassifier_class, y = class)) +
  geom_bar(stat = "identity") +
  labs(title = "Counts of type of molecule", x = "class", y = "count") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))

ggplot(noflatData_withmole, aes(x = npclassifier_superclass, y = class)) +
  geom_bar(stat = "identity") +
  labs(title = "Counts of type of molecule", x = "superclass", y = "count") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))
```
