---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
# Install and load packages

```{r}
library(dplyr) # version 1.1.4
library(ggplot2) # version 3.5.2
library(ggpubr) # version 0.6.0
library(tidyr) # version 1.3.1
library(FSA) # version 0.10.0
library(purrr) # version 1.0.4
library(stringr) # version 1.5.1
library(viridis) # version 0.6.5
library(tidyverse)
library(UpSetR)
library(ComplexUpset)
```

# Import data

```{r}
#import MElaS
data<-read.csv("MElaS_Values.csv")

#set categories as factors
data$infected_parasites=as.factor(data$infected_parasites)
data$organ=as.factor(data$organ)
data$parasite_species=as.factor(data$parasite_species)

# data from GNPS
molecules<-read.csv("DiseaseToleranceGNPSResults.csv")

# data from original feature table
metadata<-read.csv("DisTol_mergeddata.csv")
metadata<-na.omit(metadata)
ATTRIBUTE_infected_parasites<-as.factor(metadata$ATTRIBUTE_infected_parasites) 
ATTRIBUTE_endpoint<-as.factor(metadata$ATTRIBUTE_endpoint)

metadata<-filter(metadata, ATTRIBUTE_infected_parasites!='10000', ) # remove 10000
```

# Create patterns of early, late, early and late, inelastic and unchanged

```{r}
dataWithMElaSTypes <- data %>%
  mutate(
    Type = case_when(
      (MElaS_0v3v24 >= 8.60136153) & (MElaS_0v12v24 >= 7.604416279) ~ "early_late_elastic",
      MElaS_0v3v24 >= 8.60136153  ~"early_elastic",
        MElaS_0v12v24 >= 7.604416279  ~"late_elastic",
      (FC_3w <= 1.2 & FC_3w >=0.8) & (FC_12w <=1.2 & FC_12w >=0.8) & (FC_24w <=1.2 & FC_24w >=0.8) ~"unchanged",
      TRUE ~ "inelastic"
      
    )
  ) 

# Make a new dataset that removes all metabolites that were categorized "unchanged"
noflatData<-filter(dataWithMElaSTypes, Type!="unchanged", )

```

## Chemical annotation of all metabolites using GNPS data

```{r}
# Merge GNPS data with MElaS data to identify annotated molecules
dataWithMElaSTypes <- dataWithMElaSTypes %>%
  mutate(extracted_ID = str_remove_all(ID, "X(\\d+).(\\d+)_(\\d+).(\\d+)_"))

merged_all_df <- merge(
  molecules,
  dataWithMElaSTypes,
  by.x = "X.Scan.",
  by.y = "extracted_ID",
  all = FALSE 
)

merged_all_df <- merged_all_df[, c( 'ID', 'MassDiff', 'Compound_Name', 'superclass', 'class', 'MElaS_0v3v24', 'MElaS_0v12v24','organ', 'parasite_species', 'infected_parasites')]

write.csv(merged_all_df, "Annotated_MElaS_df.csv", row.names = FALSE) #export file of all features with MElaS and chemical annotation

# Merge GNPS data with MElaS data to identify annotated molecules not classified as "unchanged"
noflatData <- noflatData %>%
  mutate(extracted_ID = str_remove_all(ID, "X(\\d+).(\\d+)_(\\d+).(\\d+)_"))

merged_all_df_noflat <- merge(
  molecules,
  noflatData,
  by.x = "X.Scan.",
  by.y = "extracted_ID",
  all = FALSE 
)

merged_all_df_noflat <- merged_all_df_noflat[, c( 'ID', 'MassDiff', 'Compound_Name', 'superclass', 'class', 'MElaS_0v3v24', 'MElaS_0v12v24','organ', 'parasite_species', 'infected_parasites', "Type")]

write.csv(merged_all_df_noflat, "Annotated_MElaS_df_flatremoved.csv", row.names = FALSE) #export file of features not considered "unchanged" with MElaS and chemical annotation

```

```{r}
# Comparison of Early MElaS of each category within all metabolites (all metabolites included) wont need!!!!
dataWithMElaSTypes$Type <- factor(dataWithMElaSTypes$Type, levels=c("early_elastic", "late_elastic", "early_late_elastic", "inelastic", "unchanged"))
boxplot_cat <- ggplot(dataWithMElaSTypes, aes(x=Type, y=MElaS_0v3v24))+ geom_boxplot() + labs(title = "Boxplot Comparing MelaS of Each Category", x = "Pattern", y = "MElaS (Early)") +theme_bw() +theme(axis.text.y = element_text(size=15))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=20)) + theme(axis.title.y = element_text(size = rel(2.2))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(title=element_blank(), axis.title.x = element_blank())
print(boxplot_cat)

# Comparison of Late MElaS of each category within all metabolites (all metabolites included) wont need !!!!
dataWithMElaSTypes$Type <- factor(dataWithMElaSTypes$Type, levels=c("early_elastic", "late_elastic", "early_late_elastic", "inelastic", "unchanged"))
boxplot_cat <- ggplot(dataWithMElaSTypes, aes(x=Type, y=MElaS_0v12v24))+ geom_boxplot() + labs(title = "Boxplot Comparing MelaS of Each Category", x = "Pattern", y = "MElaS (Late)") +theme_bw() +theme(axis.text.y = element_text(size=15))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=20)) + theme(axis.title.y = element_text(size = rel(2.2))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(title=element_blank(), axis.title.x = element_blank())
print(boxplot_cat)



# Bar plot illustrating distribution of each category within metabolites. Metabolites considered flat were included in this figure
barplot_cat_flat <- ggplot(data=dataWithMElaSTypes, aes(x = Type)) + geom_bar() + labs(x="Pattern", y="Count") +theme_bw() +theme(axis.text.y = element_text(size=15)) + theme(axis.text.x = element_text(size=15)) + theme(axis.title.y = element_text(size = rel(2.2))) + theme(axis.title.x = element_text(size = rel(1.8))) + theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=20)) + theme(title=element_blank(), axis.title.x = element_blank())
print(barplot_cat_flat)

# Bar plot illustrating distribution of each category within all metabolites. Metabolites considered unchanged were removed from this data set
#dataWithMElaSTypesnoflat<-filter(dataWithMElaSTypes, Type!="unchanged", )
noflatData$Type <- factor(noflatData$Type, levels=c("early_elastic", "late_elastic", "early_late_elastic", "inelastic"))

barplot_cat_noflat <- ggplot(noflatData, aes(x = Type)) + geom_bar() + labs(x="Pattern", y="Count") +theme_bw() +theme(axis.text.y = element_text(size=15)) + theme(axis.text.x = element_text(size=15)) + theme(axis.title.y = element_text(size = rel(2.2))) + theme(axis.title.x = element_text(size = rel(2.2))) + theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=20)) + theme(title=element_blank(), axis.title.x = element_blank())
print(barplot_cat_noflat)

```

```{r}
# data frames for each strain and concentration combination
JR_1000_noflatData<- noflatData %>%
  filter(parasite_species=="JR", infected_parasites=="1000")

JR_100_noflatData<- noflatData %>%
  filter(parasite_species=="JR", infected_parasites=="100")

CL_1000_noflatData<- noflatData %>%
  filter(parasite_species=="CL", infected_parasites=="1000")

CL_100_noflatData<- noflatData %>%
  filter(parasite_species=="CL", infected_parasites=="100")
```

# Differences in non-zero MElaS values by sampling site
## Run a Benjamini–Hochberg FDR corrected Dunn’s post hoc multiple comparison test by early and late MElaS and by stain/inoculation

```{r}
#JR 1000
#early elasticity 
#Kruskal Wallis test to compare significance among organs
JR_1000_noflatData_early<- JR_1000_noflatData %>%
  filter(MElaS_0v3v24 !=0) %>%
  mutate(log_MElaS_early = log10(MElaS_0v3v24))

kruskal.test(log_MElaS_early ~ organ, data = JR_1000_noflatData_early) #organs are significantly different p<2.2e-16

 #Benjamini–Hochberg FDR corrected Dunn’s post hoc multiple comparison test
dunntestJR1000_early <- dunnTest(log_MElaS_early ~ organ,
         data=JR_1000_noflatData_early,
         method="bh")
dunn_df_JR1000_early <- dunntestJR1000_early$res

dunn_df_JR1000_early <- dunn_df_JR1000_early %>%
  separate_wider_delim(Comparison, delim = " - ", names = c("Tissue1", "Tissue2"))

tissues <- unique(c(dunn_df_JR1000_early$Tissue1, dunn_df_JR1000_early$Tissue2))

pairwise_matrix <- matrix(
  NA,
  nrow = length(tissues),
  ncol = length(tissues),
  dimnames = list(tissues, tissues)
)

for(i in 1:nrow(dunn_df_JR1000_early)) {
  
  t1 <- dunn_df_JR1000_early$Tissue1[i]
  t2 <- dunn_df_JR1000_early$Tissue2[i]
  p  <- dunn_df_JR1000_early$P.adj[i]
  
  pairwise_matrix[t1, t2] <- p
  pairwise_matrix[t2, t1] <- p   # mirror value
}

diag(pairwise_matrix) <- 1
pairwise_table <- as.data.frame(pairwise_matrix)

write.csv(pairwise_matrix, "dunn_df_JR1000_early.csv", row.names = TRUE) #export dunn's test results

#Boxplots of Log-scaled MElaS values comparisons of early MElaS values for each organ
colors <- c(
  "Duodenum"= "#8DD3C7", "Heart_A"= "#FFFFB3", "Heart_B" = "#BEBADA", "Heart_C" = "#FB8072",
  "Heart_D"= "#80B1D3","Esophagus"= "#FDB462", "Upper_Stomach"= "#B3DE69", "Illeum"= "#FCCDE5",
  "Cecum"= "#D9D9D9", "Jejunum"="#BC80BD", "Distal_Colon"= "#CCEBC5", "Proximal_Colon"= "#FFED6F"
)
boxplot_allorgans_JR1000_early <- ggplot(JR_1000_noflatData_early, aes(x = reorder(organ, -log_MElaS_early, FUN = median), y = log_MElaS_early, fill=organ))+ geom_boxplot() +scale_fill_manual(values = colors) +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=15)) + theme(axis.title.y = element_text(size = rel(1.7))) + theme(axis.title.x = element_text(size = rel(1.7))) +  labs(x="Organ", y=expression(Log [10]~MElaS~(Early))) + theme(legend.position="none")

print(boxplot_allorgans_JR1000_early)


#JR 1000
#late
#Kruskal Wallis test to compare significance among organs
JR_1000_noflatData_late<- JR_1000_noflatData %>%
  filter(MElaS_0v12v24 !=0) %>%
  mutate(log_MElaS_late = log10(MElaS_0v12v24))

kruskal.test(log_MElaS_late ~ organ, data = JR_1000_noflatData_late) #organs are significantly different p<2.2e-16

 #Benjamini–Hochberg FDR corrected Dunn’s post hoc multiple comparison test
dunntestJR1000_late <- dunnTest(log_MElaS_late ~ organ,
         data=JR_1000_noflatData_late,
         method="bh")
dunn_df_JR1000_late <- dunntestJR1000_late$res

dunn_df_JR1000_late <- dunn_df_JR1000_late %>%
  separate_wider_delim(Comparison, delim = " - ", names = c("Tissue1", "Tissue2"))

tissues <- unique(c(dunn_df_JR1000_late$Tissue1, dunn_df_JR1000_late$Tissue2))

pairwise_matrix <- matrix(
  NA,
  nrow = length(tissues),
  ncol = length(tissues),
  dimnames = list(tissues, tissues)
)

for(i in 1:nrow(dunn_df_JR1000_late)) {
  
  t1 <- dunn_df_JR1000_late$Tissue1[i]
  t2 <- dunn_df_JR1000_late$Tissue2[i]
  p  <- dunn_df_JR1000_late$P.adj[i]
  
  pairwise_matrix[t1, t2] <- p
  pairwise_matrix[t2, t1] <- p   # mirror value
}

diag(pairwise_matrix) <- 1
pairwise_table <- as.data.frame(pairwise_matrix)

write.csv(pairwise_matrix, "dunn_df_JR1000_late.csv", row.names = TRUE) #export dunn's test results

#Log-scaled MElaS values comparisons of early and late MElaS values for each organ
colors <- c(
  "Duodenum"= "#8DD3C7", "Heart_A"= "#FFFFB3", "Heart_B" = "#BEBADA", "Heart_C" = "#FB8072",
  "Heart_D"= "#80B1D3","Esophagus"= "#FDB462", "Upper_Stomach"= "#B3DE69", "Illeum"= "#FCCDE5",
  "Cecum"= "#D9D9D9", "Jejunum"="#BC80BD", "Distal_Colon"= "#CCEBC5", "Proximal_Colon"= "#FFED6F"
)
boxplot_allorgans_JR1000_late <- ggplot(JR_1000_noflatData_late, aes(x = reorder(organ, -log_MElaS_late, FUN = median), y = log_MElaS_late, fill=organ))+ geom_boxplot()+scale_fill_manual(values = colors) +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=15)) + theme(axis.title.y = element_text(size = rel(1.7))) + theme(axis.title.x = element_text(size = rel(1.7))) +  labs(x="Organ", y=expression(Log [10]~MElaS~(Late))) + theme(legend.position="none")

print(boxplot_allorgans_JR1000_late)

#CL 1000
#early
#Kruskal Wallis test to compare significance among organs
CL_1000_noflatData_early<- CL_1000_noflatData %>%
  filter(MElaS_0v3v24 !=0) %>%
  mutate(log_MElaS_early = log10(MElaS_0v3v24))

kruskal.test(log_MElaS_early ~ organ, data = CL_1000_noflatData_early) #organs are significantly different p<2.2e-16

 #Benjamini–Hochberg FDR corrected Dunn’s post hoc multiple comparison test
dunntestCL1000_early <- dunnTest(log_MElaS_early ~ organ,
         data=CL_1000_noflatData_early,
         method="bh")
dunn_df_CL1000_early <- dunntestCL1000_early$res

dunn_df_CL1000_early <- dunn_df_CL1000_early %>%
  separate_wider_delim(Comparison, delim = " - ", names = c("Tissue1", "Tissue2"))

tissues <- unique(c(dunn_df_CL1000_early$Tissue1, dunn_df_CL1000_early$Tissue2))

pairwise_matrix <- matrix(
  NA,
  nrow = length(tissues),
  ncol = length(tissues),
  dimnames = list(tissues, tissues)
)

for(i in 1:nrow(dunn_df_CL1000_early)) {
  
  t1 <- dunn_df_CL1000_early$Tissue1[i]
  t2 <- dunn_df_CL1000_early$Tissue2[i]
  p  <- dunn_df_CL1000_early$P.adj[i]
  
  pairwise_matrix[t1, t2] <- p
  pairwise_matrix[t2, t1] <- p   # mirror value
}

diag(pairwise_matrix) <- 1
pairwise_table <- as.data.frame(pairwise_matrix)

write.csv(pairwise_matrix, "dunn_df_CL1000_early.csv", row.names = TRUE) #export dunn's test results


#Log-scaled MElaS values comparisons of early and late MElaS values for each organ
colors <- c(
  "Duodenum"= "#8DD3C7", "Heart_A"= "#FFFFB3", "Heart_B" = "#BEBADA", "Heart_C" = "#FB8072",
  "Heart_D"= "#80B1D3","Esophagus"= "#FDB462", "Upper_Stomach"= "#B3DE69", "Illeum"= "#FCCDE5",
  "Cecum"= "#D9D9D9", "Jejunum"="#BC80BD", "Distal_Colon"= "#CCEBC5", "Proximal_Colon"= "#FFED6F"
)
boxplot_allorgans_CL1000_early <- ggplot(CL_1000_noflatData_early, aes(x = reorder(organ, -log_MElaS_early, FUN = median), y = log_MElaS_early, fill=organ))+ geom_boxplot()+scale_fill_manual(values = colors) +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=15)) + theme(axis.title.y = element_text(size = rel(1.7))) + theme(axis.title.x = element_text(size = rel(1.7))) +  labs(x="Organ", y=expression(Log [10]~MElaS~(Early))) + theme(legend.position="none")

print(boxplot_allorgans_CL1000_early)

#CL 1000
#late
#Kruskal Wallis test to compare significance among organs
CL_1000_noflatData_late<- CL_1000_noflatData %>%
  filter(MElaS_0v12v24 !=0) %>%
  mutate(log_MElaS_late = log10(MElaS_0v12v24))

kruskal.test(log_MElaS_late ~ organ, data = CL_1000_noflatData_late) #organs are significantly different p<2.2e-16

 #Benjamini–Hochberg FDR corrected Dunn’s post hoc multiple comparison test
dunntestCL1000_late <- dunnTest(log_MElaS_late ~ organ,
         data=CL_1000_noflatData_late,
         method="bh")
dunn_df_CL1000_late <- dunntestCL1000_late$res

dunn_df_CL1000_late <- dunn_df_CL1000_late %>%
  separate_wider_delim(Comparison, delim = " - ", names = c("Tissue1", "Tissue2"))

tissues <- unique(c(dunn_df_CL1000_late$Tissue1, dunn_df_CL1000_late$Tissue2))

pairwise_matrix <- matrix(
  NA,
  nrow = length(tissues),
  ncol = length(tissues),
  dimnames = list(tissues, tissues)
)

for(i in 1:nrow(dunn_df_CL1000_late)) {
  
  t1 <- dunn_df_CL1000_late$Tissue1[i]
  t2 <- dunn_df_CL1000_late$Tissue2[i]
  p  <- dunn_df_CL1000_late$P.adj[i]
  
  pairwise_matrix[t1, t2] <- p
  pairwise_matrix[t2, t1] <- p   # mirror value
}

diag(pairwise_matrix) <- 1
pairwise_table <- as.data.frame(pairwise_matrix)

write.csv(pairwise_matrix, "dunn_df_CL1000_late.csv", row.names = TRUE) #export dunn's test results


#Log-scaled MElaS values comparisons of early and late MElaS values for each organ
colors <- c(
  "Duodenum"= "#8DD3C7", "Heart_A"= "#FFFFB3", "Heart_B" = "#BEBADA", "Heart_C" = "#FB8072",
  "Heart_D"= "#80B1D3","Esophagus"= "#FDB462", "Upper_Stomach"= "#B3DE69", "Illeum"= "#FCCDE5", "Cecum"= "#D9D9D9", "Jejunum"="#BC80BD", "Distal_Colon"= "#CCEBC5", "Proximal_Colon"= "#FFED6F"
)
boxplot_allorgans_CL1000_late <- ggplot(CL_1000_noflatData_late, aes(x = reorder(organ, -log_MElaS_late, FUN = median), y = log_MElaS_late, fill=organ))+ geom_boxplot() +scale_fill_manual(values = colors)+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=15)) + theme(axis.title.y = element_text(size = rel(1.7))) + theme(axis.title.x = element_text(size = rel(1.7))) +  labs(x="Organ", y=expression(Log [10]~MElaS~(Late))) + theme(legend.position="none")

print(boxplot_allorgans_CL1000_late)


```

```{r}
#JR 100
#early
#Kruskal Wallis test to compare significance among organs
JR_100_noflatData_early<- JR_100_noflatData %>%
  filter(MElaS_0v3v24 !=0) %>%
  mutate(log_MElaS_early = log10(MElaS_0v3v24))

kruskal.test(log_MElaS_early ~ organ, data = JR_100_noflatData_early) #organs are significantly different p<2.2e-16

 #Benjamini–Hochberg FDR corrected Dunn’s post hoc multiple comparison test
dunntestJR100_early <- dunnTest(log_MElaS_early ~ organ,
         data=JR_100_noflatData_early,
         method="bh")
dunn_df_JR100_early <- dunntestJR100_early$res

dunn_df_JR100_early <- dunn_df_JR100_early %>%
  separate_wider_delim(Comparison, delim = " - ", names = c("Tissue1", "Tissue2"))

tissues <- unique(c(dunn_df_JR100_early$Tissue1, dunn_df_JR100_early$Tissue2))

pairwise_matrix <- matrix(
  NA,
  nrow = length(tissues),
  ncol = length(tissues),
  dimnames = list(tissues, tissues)
)

for(i in 1:nrow(dunn_df_JR100_early)) {
  
  t1 <- dunn_df_JR100_early$Tissue1[i]
  t2 <- dunn_df_JR100_early$Tissue2[i]
  p  <- dunn_df_JR100_early$P.adj[i]
  
  pairwise_matrix[t1, t2] <- p
  pairwise_matrix[t2, t1] <- p   # mirror value
}

diag(pairwise_matrix) <- 1
pairwise_table <- as.data.frame(pairwise_matrix)

write.csv(pairwise_matrix, "dunn_df_JR100_early.csv", row.names = TRUE) #export dunn's test results

#Log-scaled MElaS values comparisons of early MElaS values for each organ
colors <- c(
  "Duodenum"= "#8DD3C7", "Heart_A"= "#FFFFB3", "Heart_B" = "#BEBADA", "Heart_C" = "#FB8072",
  "Heart_D"= "#80B1D3","Esophagus"= "#FDB462", "Upper_Stomach"= "#B3DE69", "Illeum"= "#FCCDE5",
  "Cecum"= "#D9D9D9", "Jejunum"="#BC80BD", "Distal_Colon"= "#CCEBC5", "Proximal_Colon"= "#FFED6F"
)
boxplot_allorgans_JR100_early <- ggplot(JR_100_noflatData_early, aes(x = reorder(organ, -log_MElaS_early, FUN = median), y = log_MElaS_early, fill=organ))+ geom_boxplot() +scale_fill_manual(values = colors) +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=15)) + theme(axis.title.y = element_text(size = rel(1.7))) + theme(axis.title.x = element_text(size = rel(1.7))) +  labs(x="Organ", y=expression(Log [10]~MElaS~(Early))) + theme(legend.position="none")

print(boxplot_allorgans_JR100_early)

#JR 100
#late
#Kruskal Wallis test to compare significance among organs
JR_100_noflatData_late<- JR_100_noflatData %>%
  filter(MElaS_0v12v24 !=0) %>%
  mutate(log_MElaS_late = log10(MElaS_0v12v24))

kruskal.test(log_MElaS_late ~ organ, data = JR_100_noflatData_late) #organs are significantly different p<2.2e-16

 #Benjamini–Hochberg FDR corrected Dunn’s post hoc multiple comparison test
dunntestJR100_late <- dunnTest(log_MElaS_late ~ organ,
         data=JR_100_noflatData_late,
         method="bh")
dunn_df_JR100_late <- dunntestJR100_late$res

dunn_df_JR100_late <- dunn_df_JR100_late %>%
  separate_wider_delim(Comparison, delim = " - ", names = c("Tissue1", "Tissue2"))

tissues <- unique(c(dunn_df_JR100_late$Tissue1, dunn_df_JR100_late$Tissue2))

pairwise_matrix <- matrix(
  NA,
  nrow = length(tissues),
  ncol = length(tissues),
  dimnames = list(tissues, tissues)
)

for(i in 1:nrow(dunn_df_JR100_late)) {
  
  t1 <- dunn_df_JR100_late$Tissue1[i]
  t2 <- dunn_df_JR100_late$Tissue2[i]
  p  <- dunn_df_JR100_late$P.adj[i]
  
  pairwise_matrix[t1, t2] <- p
  pairwise_matrix[t2, t1] <- p   # mirror value
}

diag(pairwise_matrix) <- 1
pairwise_table <- as.data.frame(pairwise_matrix)

write.csv(pairwise_matrix, "dunn_df_JR100_late.csv", row.names = TRUE) #export dunn's test results

#Log-scaled MElaS values comparisons of early MElaS values for each organ
colors <- c(
  "Duodenum"= "#8DD3C7", "Heart_A"= "#FFFFB3", "Heart_B" = "#BEBADA", "Heart_C" = "#FB8072",
  "Heart_D"= "#80B1D3","Esophagus"= "#FDB462", "Upper_Stomach"= "#B3DE69", "Illeum"= "#FCCDE5",
  "Cecum"= "#D9D9D9", "Jejunum"="#BC80BD", "Distal_Colon"= "#CCEBC5", "Proximal_Colon"= "#FFED6F"
)
boxplot_allorgans_JR100_late <- ggplot(JR_100_noflatData_late, aes(x = reorder(organ, -log_MElaS_late, FUN = median), y = log_MElaS_late, fill=organ))+ geom_boxplot()+scale_fill_manual(values = colors) +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=15)) + theme(axis.title.y = element_text(size = rel(1.7))) + theme(axis.title.x = element_text(size = rel(1.7))) +  labs(x="Organ", y=expression(Log [10]~MElaS~(Late))) + theme(legend.position="none")

print(boxplot_allorgans_JR100_late)

#CL 100
#early
#Kruskal Wallis test to compare significance among organs
CL_100_noflatData_early<- CL_100_noflatData %>%
  filter(MElaS_0v3v24 !=0) %>%
  mutate(log_MElaS_early = log10(MElaS_0v3v24))

kruskal.test(log_MElaS_early ~ organ, data = CL_100_noflatData_early) #organs are significantly different p<2.2e-16

 #Benjamini–Hochberg FDR corrected Dunn’s post hoc multiple comparison test
dunntestCL100_early <- dunnTest(log_MElaS_early ~ organ,
         data=CL_100_noflatData_early,
         method="bh")
dunn_df_CL100_early <- dunntestCL100_early$res

dunn_df_CL100_early <- dunn_df_CL100_early %>%
  separate_wider_delim(Comparison, delim = " - ", names = c("Tissue1", "Tissue2"))

tissues <- unique(c(dunn_df_CL100_early$Tissue1, dunn_df_CL100_early$Tissue2))

pairwise_matrix <- matrix(
  NA,
  nrow = length(tissues),
  ncol = length(tissues),
  dimnames = list(tissues, tissues)
)

for(i in 1:nrow(dunn_df_CL100_early)) {
  
  t1 <- dunn_df_CL100_early$Tissue1[i]
  t2 <- dunn_df_CL100_early$Tissue2[i]
  p  <- dunn_df_CL100_early$P.adj[i]
  
  pairwise_matrix[t1, t2] <- p
  pairwise_matrix[t2, t1] <- p   # mirror value
}

diag(pairwise_matrix) <- 1
pairwise_table <- as.data.frame(pairwise_matrix)

write.csv(pairwise_matrix, "dunn_df_CL100_early.csv", row.names = TRUE) #export dunn's test results

#Log-scaled MElaS values comparisons of early MElaS values for each organ
colors <- c(
  "Duodenum"= "#8DD3C7", "Heart_A"= "#FFFFB3", "Heart_B" = "#BEBADA", "Heart_C" = "#FB8072",
  "Heart_D"= "#80B1D3","Esophagus"= "#FDB462", "Upper_Stomach"= "#B3DE69", "Illeum"= "#FCCDE5",
  "Cecum"= "#D9D9D9", "Jejunum"="#BC80BD", "Distal_Colon"= "#CCEBC5", "Proximal_Colon"= "#FFED6F"
)
boxplot_allorgans_CL100_early <- ggplot(CL_100_noflatData_early, aes(x = reorder(organ, -log_MElaS_early, FUN = median), y = log_MElaS_early, fill=organ))+ geom_boxplot()+scale_fill_manual(values = colors) +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=15)) + theme(axis.title.y = element_text(size = rel(1.7))) + theme(axis.title.x = element_text(size = rel(1.7))) +  labs(x="Organ", y=expression(Log [10]~MElaS~(Early))) + theme(legend.position="none")

print(boxplot_allorgans_CL100_early)

#CL 100
#late
#Kruskal Wallis test to compare significance among organs
CL_100_noflatData_late<- CL_100_noflatData %>%
  filter(MElaS_0v12v24 !=0) %>%
  mutate(log_MElaS_late = log10(MElaS_0v12v24))

kruskal.test(log_MElaS_late ~ organ, data = CL_100_noflatData_late) #organs are significantly different p<2.2e-16

 #Benjamini–Hochberg FDR corrected Dunn’s post hoc multiple comparison test
dunntestCL100_late <- dunnTest(log_MElaS_late ~ organ,
         data=CL_100_noflatData_late,
         method="bh")
dunn_df_CL100_late <- dunntestCL100_late$res

dunn_df_CL100_late <- dunn_df_CL100_late %>%
  separate_wider_delim(Comparison, delim = " - ", names = c("Tissue1", "Tissue2"))

tissues <- unique(c(dunn_df_CL100_late$Tissue1, dunn_df_CL100_late$Tissue2))

pairwise_matrix <- matrix(
  NA,
  nrow = length(tissues),
  ncol = length(tissues),
  dimnames = list(tissues, tissues)
)

for(i in 1:nrow(dunn_df_CL100_late)) {
  
  t1 <- dunn_df_CL100_late$Tissue1[i]
  t2 <- dunn_df_CL100_late$Tissue2[i]
  p  <- dunn_df_CL100_late$P.adj[i]
  
  pairwise_matrix[t1, t2] <- p
  pairwise_matrix[t2, t1] <- p   # mirror value
}

diag(pairwise_matrix) <- 1
pairwise_table <- as.data.frame(pairwise_matrix)

write.csv(pairwise_matrix, "dunn_df_CL100_late.csv", row.names = TRUE) #export dunn's test results

#Log-scaled MElaS values comparisons of early MElaS values for each organ
colors <- c(
  "Duodenum"= "#8DD3C7", "Heart_A"= "#FFFFB3", "Heart_B" = "#BEBADA", "Heart_C" = "#FB8072",
  "Heart_D"= "#80B1D3","Esophagus"= "#FDB462", "Upper_Stomach"= "#B3DE69", "Illeum"= "#FCCDE5", "Cecum"= "#D9D9D9", "Jejunum"="#BC80BD", "Distal_Colon"= "#CCEBC5", "Proximal_Colon"= "#FFED6F"
)
boxplot_allorgans_CL100_late <- ggplot(CL_100_noflatData_late, aes(x = reorder(organ, -log_MElaS_late, FUN = median), y = log_MElaS_late, fill=organ))+ geom_boxplot() +scale_fill_manual(values = colors)+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +theme_bw() +theme(axis.text.y = element_text(size=13))+ theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size=15)) + theme(axis.title.y = element_text(size = rel(1.7))) + theme(axis.title.x = element_text(size = rel(1.7))) +  labs(x="Organ", y=expression(Log [10]~MElaS~(Late))) + theme(legend.position="none")

print(boxplot_allorgans_CL100_late)

```

# Impact of T. cruzi strain on MElaS values
## Run a Benjamini–Hochberg FDR corrected Wilcoxon signed-rank test comparing by parasite strains
### Early MElaS
```{r}
noflatData_early<- noflatData %>%
  mutate(log_MElaS = log10(MElaS_0v3v24)) %>%
  filter(MElaS_0v3v24 !=0)

noflatData_early$Group <- paste0(noflatData_early$parasite_species, "_", noflatData_early$infected_parasites)

p_values_df <- data.frame(Organ = character(),
                          Group1 = character(),
                          Group2 = character(),
                          Test = character(),
                          P_Value = numeric(),
                          Fold_Change = numeric(),
                          stringsAsFactors = FALSE)

OrganList <- c(unique(noflatData_early$organ))
for (Organ in OrganList){
  print(Organ)
  temp <- noflatData_early[noflatData_early$organ == Organ,]
  comparisons <- combn(levels(factor(temp$Group)), 2, simplify = FALSE)
  comparisons <- c(comparisons[2], comparisons[5])
    for (comp in comparisons) {
      group1 <- comp[1]
      group2 <- comp[2]
      
      # Subset data for the two strains
      temp_subset <- temp[temp$Group %in% c(group1, group2), ]
      
      if (length(unique(temp_subset[temp_subset$Group == group1,]$log_MElaS)) > 1 & length(temp_subset[temp_subset$Group == group1,]$log_MElaS) > 2) {
        normalcy1 <- shapiro.test(temp_subset[temp_subset$Group == group1,]$log_MElaS)$p.value
        } else {
          normalcy1 <- 0 
        }
      if (length(unique(temp_subset[temp_subset$Group == group2,]$log_MElaS)) > 1 & length(temp_subset[temp_subset$Group == group2,]$log_MElaS) > 2) {
        normalcy2 <- shapiro.test(temp_subset[temp_subset$Group == group2,]$log_MElaS)$p.value
        } else {
          normalcy2 <- 0 
        }
    
      # Choose the appropriate test based on normality
      if (normalcy1 < 0.05 | normalcy2 < 0.05) {
        test_used <- "Wilcoxon"
        plotTest <- "wilcox.test"
        p_val <- wilcox.test(log_MElaS ~ Group, data = temp_subset)$p.value
      } else {
        test_used <- "T-test"
        plotTest <- "t.test"
        p_val <- t.test(log_MElaS ~ Group, data = temp_subset)$p.value
      }
      
      #make fold change
      fc <- median(temp_subset[temp_subset$Group == group1,]$log_MElaS)/median(temp_subset[temp_subset$Group == group2,]$log_MElaS)
      
      # Store results in the data frame
      p_values_df <- rbind(p_values_df, data.frame(
        Organ = Organ,
        Group1 = group1,
        Group2 = group2,
        Test = test_used,
        Fold_Change = fc,
        P_Value = p_val
      ))
    }
}
#p-value adjustment
p_values_df$p_value_adj <- 
  p.adjust(p_values_df$P_Value, method = "BH")

write.csv(p_values_df, "by_parasite_group_early_pvalues.csv", row.names = FALSE)
```

# Impact of T. cruzi strain on MElaS values
## Run a Benjamini–Hochberg FDR corrected Wilcoxon signed-rank test comparing by parasite strains
### Late MElaS
```{r}
noflatData_late<- noflatData %>%
  mutate(log_MElaS = log10(MElaS_0v12v24)) %>%
  filter(MElaS_0v12v24 !=0)

noflatData_late$Group <- paste0(noflatData_late$parasite_species, "_", noflatData_late$infected_parasites)

p_values_df <- data.frame(Organ = character(),
                          Group1 = character(),
                          Group2 = character(),
                          Test = character(),
                          P_Value = numeric(),
                          Fold_Change = numeric(),
                          stringsAsFactors = FALSE)

OrganList <- c(unique(noflatData_late$organ))
for (Organ in OrganList){
  print(Organ)
  temp <- noflatData_late[noflatData_late$organ == Organ,]
  comparisons <- combn(levels(factor(temp$Group)), 2, simplify = FALSE)
  comparisons <- c(comparisons[2], comparisons[5])
    for (comp in comparisons) {
      group1 <- comp[1]
      group2 <- comp[2]
      
      # Subset data for the two strains
      temp_subset <- temp[temp$Group %in% c(group1, group2), ]
      
      if (length(unique(temp_subset[temp_subset$Group == group1,]$log_MElaS)) > 1 & length(temp_subset[temp_subset$Group == group1,]$log_MElaS) > 2) {
        normalcy1 <- shapiro.test(temp_subset[temp_subset$Group == group1,]$log_MElaS)$p.value
        } else {
          normalcy1 <- 0 
        }
      if (length(unique(temp_subset[temp_subset$Group == group2,]$log_MElaS)) > 1 & length(temp_subset[temp_subset$Group == group2,]$log_MElaS) > 2) {
        normalcy2 <- shapiro.test(temp_subset[temp_subset$Group == group2,]$log_MElaS)$p.value
        } else {
          normalcy2 <- 0 
        }
    
      # Choose the appropriate test based on normality
      if (normalcy1 < 0.05 | normalcy2 < 0.05) {
        test_used <- "Wilcoxon"
        plotTest <- "wilcox.test"
        p_val <- wilcox.test(log_MElaS ~ Group, data = temp_subset)$p.value
      } else {
        test_used <- "T-test"
        plotTest <- "t.test"
        p_val <- t.test(log_MElaS ~ Group, data = temp_subset)$p.value
      }
      
      #make fold change
      fc <- median(temp_subset[temp_subset$Group == group1,]$log_MElaS)/median(temp_subset[temp_subset$Group == group2,]$log_MElaS)
      
      # Store results in the data frame
      p_values_df <- rbind(p_values_df, data.frame(
        Organ = Organ,
        Group1 = group1,
        Group2 = group2,
        Test = test_used,
        Fold_Change = fc,
        P_Value = p_val
      ))
    }
}
#p-value adjustment
p_values_df$p_value_adj <- 
  p.adjust(p_values_df$P_Value, method = "BH")

write.csv(p_values_df, "by_parasite_group_late_pvalues.csv", row.names = FALSE)
```

# Impact of Inoculum strain on MElaS values
## Run a Benjamini–Hochberg FDR corrected Wilcoxon signed-rank test comparing by inoculum
### Early MElaS
```{r}
noflatData_early<- noflatData %>%
  mutate(log_MElaS = log10(MElaS_0v3v24)) %>%
  filter(MElaS_0v3v24 !=0)

noflatData_early$Group <- paste0(noflatData_early$parasite_species, "_", noflatData_early$infected_parasites)

p_values_df <- data.frame(Organ = character(),
                          Group1 = character(),
                          Group2 = character(),
                          Test = character(),
                          P_Value = numeric(),
                          Fold_Change = numeric(),
                          stringsAsFactors = FALSE)

OrganList <- c(unique(noflatData_early$organ))
for (Organ in OrganList){
  print(Organ)
  temp <- noflatData_early[noflatData_early$organ == Organ,]
  comparisons <- combn(levels(factor(temp$Group)), 2, simplify = FALSE)
  comparisons <- c(comparisons[1], comparisons[6])
    for (comp in comparisons) {
      group1 <- comp[1]
      group2 <- comp[2]
      
      # Subset data for the two strains
      temp_subset <- temp[temp$Group %in% c(group1, group2), ]
      
      if (length(unique(temp_subset[temp_subset$Group == group1,]$log_MElaS)) > 1 & length(temp_subset[temp_subset$Group == group1,]$log_MElaS) > 2) {
        normalcy1 <- shapiro.test(temp_subset[temp_subset$Group == group1,]$log_MElaS)$p.value
        } else {
          normalcy1 <- 0 
        }
      if (length(unique(temp_subset[temp_subset$Group == group2,]$log_MElaS)) > 1 & length(temp_subset[temp_subset$Group == group2,]$log_MElaS) > 2) {
        normalcy2 <- shapiro.test(temp_subset[temp_subset$Group == group2,]$log_MElaS)$p.value
        } else {
          normalcy2 <- 0 
        }
    
      # Choose the appropriate test based on normality
      if (normalcy1 < 0.05 | normalcy2 < 0.05) {
        test_used <- "Wilcoxon"
        plotTest <- "wilcox.test"
        p_val <- wilcox.test(log_MElaS ~ Group, data = temp_subset)$p.value
      } else {
        test_used <- "T-test"
        plotTest <- "t.test"
        p_val <- t.test(log_MElaS ~ Group, data = temp_subset)$p.value
      }
      
      #make fold change
      fc <- median(temp_subset[temp_subset$Group == group1,]$log_MElaS)/median(temp_subset[temp_subset$Group == group2,]$log_MElaS)
      
      # Store results in the data frame
      p_values_df <- rbind(p_values_df, data.frame(
        Organ = Organ,
        Group1 = group1,
        Group2 = group2,
        Test = test_used,
        Fold_Change = fc,
        P_Value = p_val
      ))
    }
}
#p-value adjustment
p_values_df$p_value_adj <- 
  p.adjust(p_values_df$P_Value, method = "BH")

write.csv(p_values_df, "by_inoculum_group_early_pvalues.csv", row.names = FALSE)
```

# Impact of Inoculum strain on MElaS values
## Run a Benjamini–Hochberg FDR corrected Wilcoxon signed-rank test comparing by inoculum
### Late MElaS

```{r}
noflatData_late<- noflatData %>%
  mutate(log_MElaS = log10(MElaS_0v12v24)) %>%
  filter(MElaS_0v12v24 !=0)

noflatData_late$Group <- paste0(noflatData_late$parasite_species, "_", noflatData_late$infected_parasites)

p_values_df <- data.frame(Organ = character(),
                          Group1 = character(),
                          Group2 = character(),
                          Test = character(),
                          P_Value = numeric(),
                          Fold_Change = numeric(),
                          stringsAsFactors = FALSE)

OrganList <- c(unique(noflatData_late$organ))
for (Organ in OrganList){
  print(Organ)
  temp <- noflatData_late[noflatData_late$organ == Organ,]
  comparisons <- combn(levels(factor(temp$Group)), 2, simplify = FALSE)
  comparisons <- c(comparisons[1], comparisons[6])
    for (comp in comparisons) {
      group1 <- comp[1]
      group2 <- comp[2]
      
      # Subset data for the two strains
      temp_subset <- temp[temp$Group %in% c(group1, group2), ]
      
      if (length(unique(temp_subset[temp_subset$Group == group1,]$log_MElaS)) > 1 & length(temp_subset[temp_subset$Group == group1,]$log_MElaS) > 2) {
        normalcy1 <- shapiro.test(temp_subset[temp_subset$Group == group1,]$log_MElaS)$p.value
        } else {
          normalcy1 <- 0 
        }
      if (length(unique(temp_subset[temp_subset$Group == group2,]$log_MElaS)) > 1 & length(temp_subset[temp_subset$Group == group2,]$log_MElaS) > 2) {
        normalcy2 <- shapiro.test(temp_subset[temp_subset$Group == group2,]$log_MElaS)$p.value
        } else {
          normalcy2 <- 0 
        }
    
      # Choose the appropriate test based on normality
      if (normalcy1 < 0.05 | normalcy2 < 0.05) {
        test_used <- "Wilcoxon"
        plotTest <- "wilcox.test"
        p_val <- wilcox.test(log_MElaS ~ Group, data = temp_subset)$p.value
      } else {
        test_used <- "T-test"
        plotTest <- "t.test"
        p_val <- t.test(log_MElaS ~ Group, data = temp_subset)$p.value
      }
      
      #make fold change
      fc <- median(temp_subset[temp_subset$Group == group1,]$log_MElaS)/median(temp_subset[temp_subset$Group == group2,]$log_MElaS)
      
      # Store results in the data frame
      p_values_df <- rbind(p_values_df, data.frame(
        Organ = Organ,
        Group1 = group1,
        Group2 = group2,
        Test = test_used,
        Fold_Change = fc,
        P_Value = p_val
      ))
    }
}
#p-value adjustment
p_values_df$p_value_adj <- 
  p.adjust(p_values_df$P_Value, method = "BH")

write.csv(p_values_df, "by_inoculum_group_late_pvalues.csv", row.names = FALSE)
```

## Visulatization boxplots log-scaled early MElaS values by parasite strain 
###Early

```{r}
#Figure: Log-scaled MElaS values for each organ by inoculation concentration 100
# Colors
fill_colors <- c(
  "CL"  = "lightblue", 
  "JR"  = "salmon"
)

noflatData$strain_conc <- paste0(
  noflatData$parasite_species, "_",
  noflatData$infected_parasites
)
noflatData_100_Early <- noflatData %>%
  mutate(log_MElaS = log10(MElaS_0v3v24)) %>%
  filter(infected_parasites=="100")

noflatData_100_Early$organ <- factor(noflatData_100_Early$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(data=noflatData_100_Early, aes(x = as.factor(organ), y = log_MElaS, fill = parasite_species)) + geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8))  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ labs(x="Organ", y=expression(Log [10]~MElaS~(Early)))+
  
  scale_fill_manual(values = fill_colors,
                    name = "Strain") + 
  coord_cartesian(ylim = c(-3.5, 2)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 15),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.7)),
    axis.title.y = element_text(size = rel(1.7)),
    legend.text = element_text(size = 15),
    legend.title = element_text(size = rel(1.5)),
    plot.title = element_text(size = 20)
  )
#Figure: Log-scaled MElaS values for each organ by inoculation concentration 1000
# Colors
fill_colors2 <- c(
  "CL"  = "steel blue", 
  "JR"  = "red3"
)
noflatData_1000_Early <- noflatData %>%
  mutate(log_MElaS = log10(MElaS_0v3v24)) %>%
  filter(infected_parasites=="1000")

noflatData_1000_Early$organ <- factor(noflatData_1000_Early$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(data=noflatData_1000_Early, aes(x = as.factor(organ), y = log_MElaS, fill = parasite_species)) + geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ labs(x="Organ", y=expression(Log [10]~MElaS~(Early)))+
  
  scale_fill_manual(values = fill_colors2,
                    name = "Strain") + 
  coord_cartesian(ylim = c(-3.5, 2)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 15),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.7)),
    axis.title.y = element_text(size = rel(1.7)),
    legend.text = element_text(size = 15),
    legend.title = element_text(size = rel(1.5)),
    plot.title = element_text(size = 20)
  )



```

## Visulatization boxplots log-scaled late MElaS values by parasite strain 
###Late

```{r}

#Figure: Log-scaled MElaS values for each organ by inoculation concentration 100
# Colors
fill_colors <- c(
  "CL"  = "lightblue", 
  "JR"  = "salmon"
)

noflatData$strain_conc <- paste0(
  noflatData$parasite_species, "_",
  noflatData$infected_parasites
)
noflatData_100_Late <- noflatData %>%
  mutate(log_MElaS = log10(MElaS_0v12v24)) %>%
  filter(infected_parasites=="100")

noflatData_100_Late$organ <- factor(noflatData_100_Late$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(data=noflatData_100_Late, aes(x = as.factor(organ), y = log_MElaS, fill = parasite_species)) + geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8))  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ labs(x="Organ", y=expression(Log [10]~MElaS~(Late)))+
  
  scale_fill_manual(values = fill_colors,
                    name = "Strain") + 
  coord_cartesian(ylim = c(-3.5, 2)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 15),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.7)),
    axis.title.y = element_text(size = rel(1.7)),
    legend.text = element_text(size = 15),
    legend.title = element_text(size = rel(1.5)),
    plot.title = element_text(size = 20)
  )
#Figure: Log-scaled MElaS values for each organ by inoculation concentration 1000
# Colors
fill_colors2 <- c(
  "CL"  = "steel blue", 
  "JR"  = "red3"
)
noflatData_1000_Late <- noflatData %>%
  mutate(log_MElaS = log10(MElaS_0v12v24)) %>%
  filter(infected_parasites=="1000")

noflatData_1000_Late$organ <- factor(noflatData_1000_Late$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(data=noflatData_1000_Late, aes(x = as.factor(organ), y = log_MElaS, fill = parasite_species)) + geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ labs(x="Organ", y=expression(Log [10]~MElaS~(Late)))+
  
  scale_fill_manual(values = fill_colors2,
                    name = "Strain") + 
  coord_cartesian(ylim = c(-3.5, 2)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 15),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.7)),
    axis.title.y = element_text(size = rel(1.7)),
    legend.text = element_text(size = 15),
    legend.title = element_text(size = rel(1.5)),
    plot.title = element_text(size = 20)
  )


```


## Visulatization boxplots log-scaled early MElaS values by inoculum

```{r}
#Figure: Log-scaled MElaS values for each organ by CL Brener
# Colors
fill_colors <- c(
  "100"  = "lightblue", 
  "1000"  = "steel blue"
)

noflatData$strain_conc <- paste0(
  noflatData$parasite_species, "_",
  noflatData$infected_parasites
)
noflatData_100_Early <- noflatData %>%
  mutate(log_MElaS = log10(MElaS_0v3v24)) %>%
  filter(parasite_species=="CL")

noflatData_100_Early$organ <- factor(noflatData_100_Early$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(data=noflatData_100_Early, aes(x = as.factor(organ), y = log_MElaS, fill = infected_parasites)) + geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) + stat_compare_means(
    aes(group = infected_parasites),
    method = "wilcox.test",
    label = "p.signif",
    hide.ns = TRUE,
    p.adjust.method = "BH",
    bracket.size = 0.8,
    label.y = 1.8
  )  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ labs(x="Organ", y=expression(Log [10]~MElaS~(Early)))+
  
  scale_fill_manual(values = fill_colors,
                    name = "Inoculum") + 
  coord_cartesian(ylim = c(-3.5, 2)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 15),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.7)),
    axis.title.y = element_text(size = rel(1.7)),
    legend.text = element_text(size = 15),
    legend.title = element_text(size = rel(1.5)),
    plot.title = element_text(size = 20)
  )
#Figure: Log-scaled MElaS values for each organ by inoculation concentration JR
# Colors
fill_colors2 <- c(
  "100"  = "salmon", 
  "1000"  = "red3"
)
noflatData_1000_Early <- noflatData %>%
  mutate(log_MElaS = log10(MElaS_0v3v24)) %>%
  filter(parasite_species=="JR")

noflatData_1000_Early$organ <- factor(noflatData_1000_Early$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(data=noflatData_1000_Early, aes(x = as.factor(organ), y = log_MElaS, fill = infected_parasites)) + geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) + stat_compare_means(
    aes(group = infected_parasites),
    method = "wilcox.test",
    label = "p.signif",
    hide.ns = TRUE,
    p.adjust.method = "BH",
    bracket.size = 0.8,
    label.y = 1.8
  )  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ labs(x="Organ", y=expression(Log [10]~MElaS~(Early)))+
  
  scale_fill_manual(values = fill_colors2,
                    name = "inoculum") + 
  coord_cartesian(ylim = c(-3.5, 2)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 15),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.7)),
    axis.title.y = element_text(size = rel(1.7)),
    legend.text = element_text(size = 15),
    legend.title = element_text(size = rel(1.5)),
    plot.title = element_text(size = 20)
  )



```

## Visulatization boxplots log-scaled late MElaS values by inoculum

```{r}

#Figure: Log-scaled MElaS values for each organ by CL Brener
# Colors
fill_colors <- c(
  "100"  = "lightblue", 
  "1000"  = "steel blue"
)

noflatData$strain_conc <- paste0(
  noflatData$parasite_species, "_",
  noflatData$infected_parasites
)
noflatData_100_Late <- noflatData %>%
  mutate(log_MElaS = log10(MElaS_0v12v24)) %>%
  filter(parasite_species=="CL")

noflatData_100_Late$organ <- factor(noflatData_100_Late$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(data=noflatData_100_Late, aes(x = as.factor(organ), y = log_MElaS, fill = infected_parasites)) + geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) + stat_compare_means(
    aes(group = infected_parasites),
    method = "wilcox.test",
    label = "p.signif",
    hide.ns = TRUE,
    p.adjust.method = "BH",
    bracket.size = 0.8,
    label.y = 1.8
  )  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ labs(x="Organ", y=expression(Log [10]~MElaS~(Late)))+
  
  scale_fill_manual(values = fill_colors,
                    name = "Inoculum") + 
  coord_cartesian(ylim = c(-3.5, 2)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 15),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.7)),
    axis.title.y = element_text(size = rel(1.7)),
    legend.text = element_text(size = 15),
    legend.title = element_text(size = rel(1.5)),
    plot.title = element_text(size = 20)
  )
#Figure: Log-scaled MElaS values for each organ by inoculation concentration 1000
# Colors
fill_colors2 <- c(
  "100"  = "salmon", 
  "1000"  = "red3"
)
noflatData_1000_Late <- noflatData %>%
  mutate(log_MElaS = log10(MElaS_0v12v24)) %>%
  filter(parasite_species=="JR")

noflatData_1000_Late$organ <- factor(noflatData_1000_Late$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

ggplot(data=noflatData_1000_Late, aes(x = as.factor(organ), y = log_MElaS, fill = infected_parasites)) + geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) + stat_compare_means(
    aes(group = infected_parasites),
    method = "wilcox.test",
    label = "p.signif",
    hide.ns = TRUE,
    p.adjust.method = "BH",
    bracket.size = 0.8,
    label.y = 1.8
  )  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ labs(x="Organ", y=expression(Log [10]~MElaS~(Late)))+
  
  scale_fill_manual(values = fill_colors2,
                    name = "Inoculum") + 
  coord_cartesian(ylim = c(-3.5, 2)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 15),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.7)),
    axis.title.y = element_text(size = rel(1.7)),
    legend.text = element_text(size = 15),
    legend.title = element_text(size = rel(1.5)),
    plot.title = element_text(size = 20)
  )


```
redo!!!!
# Impact of T. cruzi strain on the proportion of elastic metabolites
##Early
```{r}
#percentage of nonzeros by organ 
proportion_non_zero_by_organ_early <- noflatData %>%
  group_by(organ, parasite_species, infected_parasites) %>%
  summarise(
    non_zero_count_organ=sum(Type == "inelastic"),
    total_rows_organ=n(),
    zero_count=total_rows_organ-non_zero_count_organ,
    proportion_non_zero_organ = (non_zero_count_organ / total_rows_organ) *100,
    proportion_non_zero_organ =format(round(proportion_non_zero_organ, 3), nsmall = 3)
  )

proportion_non_zero_by_organ_early$proportion_non_zero_organ=as.numeric(proportion_non_zero_by_organ_early$proportion_non_zero_organ)
```

## Percentage of Non-Zero MElas Scores per Organ Barplots

```{r}

proportion_non_zero_by_organ_early$strain_conc <- paste0(
  proportion_non_zero_by_organ_early$parasite_species, "_",
  proportion_non_zero_by_organ_early$infected_parasites
)
proportion_non_zero_by_organ_early$organ <- factor(proportion_non_zero_by_organ_early$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

proportion_non_zero_by_organ_early_1000 <- proportion_non_zero_by_organ_early %>%
  filter(infected_parasites=="1000")

proportion_non_zero_by_organ_early_100 <- proportion_non_zero_by_organ_early %>%
  filter(infected_parasites=="100")


#Figure 5A. Comparison of the proportion of metabolites with a MElaS value of nonzero for 1000
# Colors
fill_colors <- c(
  "CL" = "steel blue",
  "JR" = "red3"
)

ggplot(data = proportion_non_zero_by_organ_early_1000,
       aes(x = as.factor(organ),
           y = proportion_non_zero_organ,
           fill = parasite_species)) +
  geom_col(position = position_dodge(width = 0.9)) +
  
  scale_fill_manual(values = fill_colors,
                    name = "Strain") +
  
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.8)),
    axis.title.y = element_text(size = rel(1.8)),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = rel(1.3))
  ) +
  labs(x = "Organ", y = "Percentage (%)")

#Figure 5B. Comparison of the proportion of metabolites with a MElaS value of nonzero for each organ by 100
# Colors
fill_colors <- c(
  "CL"  = "lightblue", 
  "JR"  = "salmon"
)

ggplot(data = proportion_non_zero_by_organ_early_100,
       aes(x = as.factor(organ),
           y = proportion_non_zero_organ,
           fill = parasite_species)) +
  geom_col(position = position_dodge(width = 0.9)) +
  
  scale_fill_manual(values = fill_colors,
                    name = "Strain") +
  
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.8)),
    axis.title.y = element_text(size = rel(1.8)),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = rel(1.3))
  ) +
  labs(x = "Organ", y = "Percentage (%)")


```

# Impact of inoculation concentration and T. cruzi strain on the proportion of elastic metabolites
##Late
```{r}
#percentage of nonzeros by organ 
proportion_non_zero_by_organ_late <- dataWithMElaSTypes %>%
  group_by(organ, parasite_species, infected_parasites) %>%
  summarise(
    non_zero_count_organ=sum(Type == "late_elastic"),
    total_rows_organ=n(),
    proportion_non_zero_organ = (non_zero_count_organ / total_rows_organ) *100,
    proportion_non_zero_organ =format(round(proportion_non_zero_organ, 3), nsmall = 3),
    zero_count=total_rows_organ-non_zero_count_organ
  )

proportion_non_zero_by_organ_late$proportion_non_zero_organ=as.numeric(proportion_non_zero_by_organ_late$proportion_non_zero_organ)
```

## Percentage of Non-Zero MElas Scores per Organ Barplots

```{r}

proportion_non_zero_by_organ_late$strain_conc <- paste0(
  proportion_non_zero_by_organ_late$parasite_species, "_",
  proportion_non_zero_by_organ_late$infected_parasites
)
proportion_non_zero_by_organ_late$organ <- factor(proportion_non_zero_by_organ_late$organ, levels=c("Heart_A", "Heart_B", "Heart_C", "Heart_D", "Esophagus", "Upper_Stomach", "Duodenum", "Jejunum", "Ileum", "Cecum", "Proximal_Colon", "Distal_Colon"))

proportion_non_zero_by_organ_late_1000 <- proportion_non_zero_by_organ_late %>%
  filter(infected_parasites=="1000")

proportion_non_zero_by_organ_late_100 <- proportion_non_zero_by_organ_late %>%
  filter(infected_parasites=="100")


#Figure 5A. Comparison of the proportion of metabolites with a MElaS value of nonzero for 1000
# Colors
fill_colors <- c(
  "CL" = "steel blue",
  "JR" = "red3"
)

ggplot(data = proportion_non_zero_by_organ_late_1000,
       aes(x = as.factor(organ),
           y = proportion_non_zero_organ,
           fill = parasite_species)) +
  geom_col(position = position_dodge(width = 0.9)) +
  
  scale_fill_manual(values = fill_colors,
                    name = "Strain") +
  
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.8)),
    axis.title.y = element_text(size = rel(1.8)),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = rel(1.3))
  ) +
  labs(x = "Organ", y = "Percentage (%)")

#Figure 5B. Comparison of the proportion of metabolites with a MElaS value of nonzero for each organ by 100
# Colors
fill_colors <- c(
  "CL"  = "lightblue", 
  "JR"  = "salmon"
)

ggplot(data = proportion_non_zero_by_organ_late_100,
       aes(x = as.factor(organ),
           y = proportion_non_zero_organ,
           fill = parasite_species)) +
  geom_col(position = position_dodge(width = 0.9)) +
  
  scale_fill_manual(values = fill_colors,
                    name = "Strain") +
  
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = rel(1.8)),
    axis.title.y = element_text(size = rel(1.8)),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = rel(1.3))
  ) +
  labs(x = "Organ", y = "Percentage (%)")


```

# UpSet Plot
```{r}
SpeciesList <- c(unique(merged_all_df_noflat$parasite_species))
TypeList <- c(unique(merged_all_df_noflat$Type))

for (species in SpeciesList){
  for (type in TypeList){
    print(species)
    print(type)
    JR_1000_noflatData_early<- merged_all_df_noflat %>%
  filter(parasite_species==species, infected_parasites=="1000", Type==type)
    upset_data <- JR_1000_noflatData_early %>%
  distinct(ID, superclass, organ) %>%
  mutate(value = 1) %>%
  pivot_wider(
    names_from = organ,
    values_from = value,
    values_fill = FALSE
  )

organ_cols <- setdiff(colnames(upset_data), c("ID", "superclass"))

# Remove ID column
plot_matrix <- as.data.frame(upset_data[, -1])

plot <- upset(
  upset_data,
  intersect = organ_cols,
  
  base_annotations = list(
    "Intersection Size" =
      intersection_size(
        mapping = aes(fill = superclass)
      )
  )
) +
  scale_fill_brewer(palette = "Set3") +
  theme_bw()
print(plot)
}}
  


```



```{r}
#by strain
merged_all_df_noflat$strain_conc <- paste0(
  merged_all_df_noflat$parasite_species, "_",
  merged_all_df_noflat$infected_parasites
)

OrganList <- c(unique(merged_all_df_noflat$organ))

for (Organ in OrganList){
  print(Organ)
  JR_1000_noflatData_early<- merged_all_df_noflat %>%
  filter(Organ==organ)
  upset_data <- JR_1000_noflatData_early %>%
  distinct(ID, superclass, strain_conc) %>%
  mutate(value = 1) %>%
  pivot_wider(
    names_from = strain_conc,
    values_from = value,
    values_fill = FALSE
  )

organ_cols <- setdiff(colnames(upset_data), c("ID", "superclass"))

# Remove ID column
plot_matrix <- as.data.frame(upset_data[, -1])

plot <- upset(
  upset_data,
  intersect = organ_cols,
  
  base_annotations = list(
    "Intersection Size" =
      intersection_size(
        mapping = aes(fill = superclass)
      )
  )
) +
  scale_fill_brewer(palette = "Set3") +
  ggtitle(paste0(Organ)) +
  theme_bw()
  print(plot)
}


```

```{r}
# by type of elasticity
merged_all_df_noflat$strain_conc <- paste0(
  merged_all_df_noflat$parasite_species, "_",
  merged_all_df_noflat$infected_parasites
)

OrganList <- c(unique(merged_all_df_noflat$organ))

StrainList <- c(unique(merged_all_df_noflat$strain_conc))

for (Organ in OrganList){
  for (Strain in StrainList){
    print(Organ)
    print(Strain)
    JR_1000_noflatData_early<- merged_all_df_noflat %>%
      filter(Organ==organ, Strain==strain_conc)
    upset_data <- JR_1000_noflatData_early %>%
      distinct(ID, superclass, Type) %>%
      mutate(value = 1) %>%
      pivot_wider(
        names_from = Type,
        values_from = value,
        values_fill = FALSE)

    organ_cols <- setdiff(colnames(upset_data), c("ID", "superclass"))

    # Remove ID column
    plot_matrix <- as.data.frame(upset_data[, -1])

    plot <- upset(
      upset_data,
      intersect = organ_cols,
      
      base_annotations = list(
        "Intersection Size" =
          intersection_size(
            mapping = aes(fill = superclass)
          )
      )
    ) +
      scale_fill_brewer(palette = "Set3") +
      ggtitle(paste0(Organ, ", ", Strain)) +
      theme_bw()
      print(plot)
  }
  }

  print(Organ)
  JR_1000_noflatData_early<- merged_all_df_noflat %>%
  filter(Organ==organ)
  upset_data <- JR_1000_noflatData_early %>%
  distinct(ID, superclass, Type) %>%
  mutate(value = 1) %>%
  pivot_wider(
    names_from = Type,
    values_from = value,
    values_fill = FALSE
  )

organ_cols <- setdiff(colnames(upset_data), c("ID", "superclass"))

# Remove ID column
plot_matrix <- as.data.frame(upset_data[, -1])

plot <- upset(
  upset_data,
  intersect = organ_cols,
  
  base_annotations = list(
    "Intersection Size" =
      intersection_size(
        mapping = aes(fill = superclass)
      )
  )
) +
  scale_fill_brewer(palette = "Set3") +
  ggtitle(paste0(Organ)) +
  theme_bw()
  print(plot)

```

